### 2020-01-23 paper_rl/neural_network.tex
Add neural_network overview

\documentclass[crop]{standalone}
\usepackage{libertine}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,backgrounds}
\usepackage{amsmath}
\begin{document}
\newcommand{\nodedistance}{1.2}%
\newcommand{\numUnits}{128}%
%\pgfmathsetmacro{\shortside}{sqrt((\nodedistance * \nodedistance ) /2)}%
\pgfmathsetmacro{\shortside}{sqrt(2)/2 * \nodedistance}%
%\pgfmathsetmacro{\solutionx}{ \shortside }%
%\pgfmathsetmacro{\solutiony}{ - (3 * \shortside + 3 * \nodedistance ) }%
%\pgfmathsetmacro{\solutiondurationx}{ \shortside }%
%\pgfmathsetmacro{\solutiondurationy}{ - (3 * \shortside + 1 * \nodedistance ) }%
\tikzstyle{texto} = [above, text width=6em, text centered]%
\newcommand{\backgroundna}[3]{%%
  \begin{pgfonlayer}{background}%
    \path (#1.south -| #1.west)+(-0.5,-0.5) node (a1) {};%
    \path (#2.north -| #2.east)+(+0.5,+0.5) node (a2) {};%
    \path[rounded corners, draw=black!50]%
      (a1) rectangle (a2);%
    \path (a1.west |- a2.north)+(0.25,-0.25) node (u1)[anchor=north west]%
      {\textit{#3}};%
  \end{pgfonlayer}}%
\newcommand{\backgroundnc}[4]{%%
  \begin{pgfonlayer}{background}%
    \path (#1.south -| #4.west)+(-0.5,-0.5) node (a1) {};%
    \path (#2.north -| #2.east)+(+0.5,+0.5) node (a2) {};%
    \path[rounded corners, draw=black!50]%
      (a1) rectangle (a2);%
    \path (a1.east |- a1.south)+(2.5,0.25) node (u1)[anchor=south west]%
      {\textit{#3}};%
  \end{pgfonlayer}}%
\newcommand{\backgroundnclass}[3]{%%
  \begin{pgfonlayer}{background}%
    \path (#1.south -| #1.west)+(-0.5,-0.8) node (a1) {};%
    \path (#2.north -| #2.east)+(+0.5,+0.5) node (a2) {};%
    \path[rounded corners, draw=black!50]%
      (a1) rectangle (a2);%
    \path (a2.west |- a1.north)+(0.,-0.) node (u1)[anchor=south east]%
      {\textit{#3}};%
  \end{pgfonlayer}}%
  \tikzstyle{place}=[circle,thick,draw=blue!75,fill=blue!20,minimum size=6mm]%
  \tikzstyle{input place}=[circle,thick,draw=olive!75,fill=olive!20,minimum size=6mm]%
  \tikzstyle{output place}=[circle,thick,draw=magenta!75,fill=magenta!20,minimum size=6mm]%
   \tikzstyle{lstm place}=[circle,thick,draw=red!75,fill=red!20,minimum size=6mm]%
   \tikzstyle{softplus place}=[circle,thick,draw=green!75,fill=green!20,minimum size=6mm]%
  \tikzstyle{transition}=[rectangle,thick,draw=black!75,%
  			  fill=black!20,minimum size=4mm]%
\begin{tikzpicture}[every label/.append style={font=\tiny}, node distance=\nodedistance cm, >=stealth',bend angle=45]
\begin{scope}
\node [input place] (s1) [label={[align=center]left: State (8)}] {};
\node [place] (s2) [above right of=s1, label={[align=center]above: Linear}] {};
\path[->] (s1) edge node {} (s2);
\node [lstm place] (s3) [right of=s2, label={[align=center]below: LSTM (\numUnits{})}] {};
\path[->] (s2) edge node {} (s3); 
\path[->] (s3) edge [loop above] node {} (s3);   
\node [lstm place] (s4) [right of=s3, label={[align=center]below: LSTM (\numUnits{})}] {};    
\path[->] (s3) edge node {} (s4);
\path[->] (s4) edge [loop above] node {} (s4);
\node [lstm place] (s9) [right of=s4, label={[align=center]below: LSTM (\numUnits{})}] {};    
\path[->] (s4) edge node {} (s9);
\path[->] (s9) edge [loop above] node {} (s9);   
\node [place] (s7) [right of=s9, label={[align=center]below:{ Linear}}] {};    
\path[->] (s9) edge node {} (s7);    
\node [output place] (s8) [right of=s7, label={[align=center]below: $\sigma$~(1)}] {};    
\path[->] (s7) edge node {} (s8);
\node [place] (s5) [above of=s7, label={[align=center]above: Linear}] {};
\path[->] (s9) edge node {} (s5);
\node [output place] (s6) [right of=s5, label={[align=center]above: $\mu$ (1)}] {};        
\backgroundna{s2}{s6}{Actor network}
\path[->] (s5) edge node {} (s6);        
\node [place] (s10) [below right of=s1, label={[align=center]below: Linear}] {};
\path[->] (s1) edge node {} (s10);    
\node [lstm place] (s11) [right of=s10, label={[align=center]above: LSTM (\numUnits{})}] {};
\path[->] (s10) edge node {} (s11);
\path[->] (s11) edge [loop below] node {} (s11);
\node [lstm place] (s12) [right of=s11, label={[align=center]above: LSTM (\numUnits{})}] {};
\path[->] (s11) edge node {} (s12);
\path[->] (s12) edge [loop below] node {} (s12);    
\node [lstm place] (s21) [right of=s12, label={[align=center]above: LSTM (\numUnits{})}] {};
\path[->] (s12) edge node {} (s21);
\path[->] (s21) edge [loop below] node {} (s21);    
\node [place] (s13) [right of=s21, label={[align=center]above: Linear}] {};
\path[->] (s21) edge node {} (s13);
\node [output place] (s17) [right of=s13, label={[align=center]above: $v_\text{classification}$ (1)}] {};
\path[->] (s13) edge node {} (s17);
\node [place] (s16) [below of=s13, label={[align=center]below: Linear}] {};
\path[->] (s21) edge node {} (s16);    
\node [output place] (s20) [right of=s16, label={[align=center]below: $v_\text{sparsity}$ (1)}] {};    
\path[->] (s16) edge node {} (s20);
\backgroundnc{s20}{s17}{Critic network}{s10}
\node [place] (s30) at (\shortside,-3*\shortside-\nodedistance) [label={[align=center]below: Linear}] {};
\path[->] (s1) edge node {} (s30);
\node [lstm place] (s31) [right of=s30, label={[align=center]above: LSTM (\numUnits{})}] {};
\path[->] (s30) edge node {} (s31);
\path[->] (s31) edge [loop below] node {} (s31);
\node [lstm place] (s32) [right of=s31, label={[align=center]above: LSTM (\numUnits{})}] {};
\path[->] (s31) edge node {} (s32);
\path[->] (s32) edge [loop below] node {} (s32);    
\node [lstm place] (s33) [right of=s32, label={[align=center]above: LSTM (\numUnits{})}] {};
\path[->] (s32) edge node {} (s33);
\path[->] (s33) edge [loop below] node {} (s33);  
\node [place] (s34) [right of=s33, label={[align=center]above: Linear}] {};
\path[->] (s33) edge node {} (s34);    
\node [output place] (s35) [right of=s34, label={[align=center]above: confidence (1)}] {};    
\path[->] (s34) edge node {} (s35);
\backgroundnclass{s30}{s35}{Classifier};
\end{scope}
\end{tikzpicture}
\end{document}