### 2020-02-03 paper_rl/paper.tex
Intro rewording + minor changes

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[conference]{IEEEtran}
%% IEEE CNS addition:
\makeatletter
\def\ps@headings{%
\def\@oddhead{\mbox{}\scriptsize\rightmark \hfil \thepage}%
\def\@evenhead{\scriptsize\thepage \hfil \leftmark\mbox{}}%
\def\@oddfoot{}%
\def\@evenfoot{}}
\makeatother
\pagestyle{empty}

\usepackage[utf8]{inputenc}
\usepackage{multirow}
\usepackage[inline]{enumitem}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{subfig}

\usepackage[nomain, toc, acronym]{glossaries}
\glsdisablehyper

\newcommand\note[2]{{\color{#1}#2}}
\newcommand\todo[1]{{\note{red}{TODO: #1}}}

% Adds a "given" symbol (vertical bar) that rescales in height (like \left and \right)
\usepackage{mathtools}
\newcommand\givenbase[1][]{\:#1\lvert\:}
\let\given\givenbase
\newcommand\sgiven{\givenbase[\delimsize]}
\DeclarePairedDelimiterX\Basics[1](){\let\given\sgiven #1}
\newcommand\Average{E\Basics}

\usepackage{tabularx}

% Macro to make entire row bold from https://tex.stackexchange.com/questions/309833/format-whole-row-of-table-as-bold
\newcommand\setrow[1]{\gdef\rowmac{#1}#1\ignorespaces}
\newcommand\clearrow{\global\let\rowmac\relax}
\clearrow

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{SparseIDS: Learning Packet Sampling with Reinforcement Learning}

\author{\IEEEauthorblockN{Maximilian Bachl, Fares Meghdouri, Joachim Fabini, Tanja Zseby}
\IEEEauthorblockA{Technische Universität Wien\\
firstname.lastname@tuwien.ac.at}}

%\date{\today}

% \IEEEoverridecommandlockouts
% \IEEEpubid{\begin{minipage}[t]{\textwidth}\ \\[10pt]
%         \centering\normalsize{xxx-x-xxxx-xxxx-x/xx/\$31.00 \copyright 2018 IEEE}
% \end{minipage}}

% \renewcommand*{\bibfont}{\footnotesize}

\maketitle%

%\thispagestyle{plain}
%\pagestyle{plain}

\newacronym{ml}{ML}{Machine Learning}
\newacronym{dl}{DL}{Deep Learning}
\newacronym{aml}{AML}{Adversarial Machine Learning}
\newacronym{ids}{IDS}{Intrusion Detection System}
\newacronym{rnn}{RNN}{Recurrent Neural Network}
\newacronym{fgsm}{FGSM}{Fast Gradient Sign Method}
\newacronym{cw}{CW}{Carlini-Wagner method}
\newacronym{pgd}{PGD}{Projected Gradient Descent}
\newacronym{pdp}{PDP}{Partial Dependence Plot}
\newacronym{ars}{ARS}{Adversarial Robustness Score}
\newacronym{ttl}{TTL}{Time-to-Live}
\newacronym{dos}{DoS}{Denial-of-Service}
\newacronym{iat}{IAT}{Interarrival time}
\newacronym{rl}{RL}{Reinforcement Learning}
\newacronym{dpi}{DPI}{Deep Packet Inspection}
\newacronym{ac}{AC}{Actor-Critic}

\newcommand{\ours}{SparseIDS}

\begin{abstract}

\glspl{rnn} have been shown to be valuable for constructing \glspl{ids} for network data. They allow analyzing network flows and already determining if a flow is malicious or not before it is over, making it possible to take action immediately. However, considering the large number of packets that have to be inspected, the question of computational as well as energy efficiency arises. We show that by using a novel \gls{rl}-based approach called \textit{\ours{}}, we can reduce the number of consumed packets by more than three fourths while keeping classification accuracy high. Comparing to various other sampling techniques, \ours{} consistently achieves higher classification accuracy with the same number of sampled packets. A major novelty of our \gls{rl}-based approach is that it can not only skip up to a predefined maximum number of packets like current approaches but can possibly skip arbitrarily many packets, saving even more computational resources for long sequences. Finally we build an automatic steering mechanism that can guide \ours{} to achieve a desired level of sparsity.

\end{abstract}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
%\begin{abstract}
%  A clear and well-documented \LaTeX\ document is presented as an
%  article formatted for publication by ACM in a conference proceedings
%  or journal publication. Based on the ``acmart'' document class, this
%  article presents and explains many of the common variations, as well
%  as many of the formatting elements an author may use in the
%  preparation of the documentation of their work.
%\end{abstract}

\maketitle

\section{Introduction}
\label{sec:introduction}

%There is a significant body of scientific work focusing on the detection of unwanted behavior in networks. In the past, a viable way of performing intrusion detection was to inspect the content of packets themselves and detect if a packet delivers potentially harmful content. More recently, with the increasing deployment of encryption, the focus now lies on features that are always available to network monitoring equipment like port numbers, protocol flags or packet sizes when encrypting above the transport layer.

% Fares:
As the need for security mechanisms in today's infrastructure communication networks grows, a substantial body of scientific work focuses on developing reliable, explainable frameworks for the identification of unwanted behaviors. In the past, a viable way to detect intrusion was to analyze the contents of the packets themselves and determine, for example, whether a packet contains potentially harmful content by matching patterns. More recently, with the increasing deployment of encryption, the focus is now on features that are always available for network monitoring devices such as port numbers, protocol flags and packet sizes when encrypting above the transport layer.

%Network communication is usually aggregated into \textit{flows}, which are commonly defined as a sequence of packets sharing certain properties. When analyzing flows, not only the aforementioned features are available but also features related to the timing of the individual packets. Various approaches have been proposed to extract features from flows and then perform anomaly detection with the extracted flows \cite{meghdouri_analysis_2018}.
%While these approaches frequently work well, it is problematic that the whole flow has to be received first and only afterwards anomaly detection is applied, revealing attack flows. Thus, we design a network \gls{ids} that operates on a per-packet basis and decides if a packet is anomalous based on features that are available even for traffic that is encrypted above the transport layer, like for example TLS or QUIC.
%At the same time, an \gls{rnn}-based \gls{ids} has the benefit of avoiding tedious feature engineering procedures, which derive statistical measures from the sequence of packet features, but providing any available information to the classifier.
%\cite{hartl_explainability_2019} showed that such a system has similar performance to other flow-based anomaly detection systems but can detect anomalies \textit{before} the flow terminates and analyzed its security with respect to adversaries extensively.

% Fares:
A network communication is typically aggregated into \textit{flows}, which are commonly defined as a sequence of packets that share certain properties. When analyzing flows, not only the aforementioned features are available but also features representing the temporal behavior of the individual packets. Various approaches have been proposed to extract flows features after which anomaly detection can be performed, in \cite{meghdouri_analysis_2018} authors compare such approaches.
While these approaches often work well, it is troublesome because, in the worst case, the whole flow must be received first and only then anomaly detection can be applied; revealing malicious flows. Thus, we design a network \gls{ids} that operates on a per-packet basis and determines whether packet is anomalous based on features that are available even for traffic that is encrypted above the transport layer, like for example TLS or QUIC.
At the same time, an \gls{rnn}-based \gls{ids} has the benefit of avoiding tedious feature engineering procedures, which derive statistical measures from the sequence of packet features, but providing any available information to the classifier.
In \cite{hartl_explainability_2019}, authors show that such an architecture has similar performance to other flow-based anomaly detection systems, but can still detect anomalies \textit{before} the flow ends and analyzed its robustness with respect to adversaries extensively.

%However, for practical use, not only classification accuracy and security are important but also computational efficiency. This follows from the fact that a rationally-acting company will only deploy an \gls{ids} if the cost that is saved by avoiding security breaches is higher than the cost of the \gls{ids} itself. The cost of the \gls{ids} can be split in the cost of purchase/installation, the cost of operating it and the cost of maintaining it. The goal we pursue here is to develop sampling strategies that do not significantly degrade classification performance while being able to only process a small fraction of the original data. Specifically, the sampling technique should

% Fares:
Nevertheless, not only the accuracy and reliability of classification, but also the computational efficiency, are critical for practical use. This follows from the fact that a rationally-acting organization will only implement an \gls{ids} if the cost saved by preventing security breaches is higher than the cost of the \gls{ids} itself. Noteworthy that the cost of an \gls{ids} can be divided into (a) the cost of purchase/installation, (b) the cost of operation and (c) the cost of maintenance.  The goal we pursue here is to develop autonomous sampling strategies that do not significantly degrade classification performance while being able to only process a small fraction of the original data. Specifically, the sampling technique should:

\begin{enumerate}
%\item \textbf{choose optimally}, taking only the data that contain the most information and skip the less relevant ones.
% Fares:
\item \textbf{choose optimally}, taking only samples that contain the most information and skip the less relevant ones.
\item have a parameter that allows to \textbf{trade off} classification performance for sparsity.
\item be \textbf{independent of the classifier} so that any classifier can be used and that the classifier doesn't have to be aware of the sampling strategy.
\item be \textbf{retrainable} so that the sampling can be continuously adapted depending on the current threat landscape.
\item be able to \textbf{skip arbitrarily} many packets since network flows can be very long but only the first couple of packets are needed to decide whether an attack occurred or not and the remaining packets can be skipped altogether.
\end{enumerate}

%For this purpose, we develop the \gls{rl}-based \gls{ids} \textit{\ours{}} fulfills the above properties. We show that a significant number of packets can be skipped while accuracy doesn't drop significantly. Compared to other common sampling strategies, \ours{} performs better when being trained on the same number of packets.

% Fares:
For this purpose we develop a \gls{rl}-based \gls{ids}: \textit{\ours{}} that fulfills the above properties. We show that a significant number of packets can be skipped while the accuracy does not drop
considerably. Compared to other common sampling techniques, \ours{} performs better when being trained on the same number of packets.

%As \ours{} uses a parameter to trade-off accuracy for sparsity, the resulting accuracy and sparsity are determined by this parameter. However, since it also might be desired to specify the sparsity itself, we propose a steering system which takes a desired sparsity as its input and modifies the parameter of the reinforcement learning in a control loop until the specified sparsity is reached.

% Fares:
The accuracy-sparsity trade-off of \ours{} can be regulated by a reinforcement learning parameter that directly affects both later quantities. However, in certain situations, a certain sparsity is to be achieved, hence we further develop and propose a steering system that fluctuates such a trade-off parameter in order to obtain a coveted sparsity in a closed control loop.

To encourage reproducibility and facilitate experimentation, we publicly release the source code, the data and the figures of this work\footnote{\url{https://github.com/CN-TU/adversarial-recurrent-ids/tree/rl}}.

\section{Related Work}

\cite{hartl_explainability_2019} build an LSTM-based recurrent classifier for Network Intrusion Detection and develop methods for explainability and to assess the security of recurrent classifiers.

\cite{yu_learning_2017} propose using \gls{rl} to let an LSTM learn to skim text. Their technique does not aim to maximize sparsity but only lets the \gls{rl} optimize classification performance for the sequence as a whole. Thus, if it learns to skip elements it is only because it is better for achieving good classification performance, not because it wants to increase sparsity.

\cite{campos_skip_2018} develop a technique that includes a \textit{skip gate} into LSTM or GRU \todo{citation needed} cells. It works without \gls{rl} but on the downside, their method can only be trained once and not be adapted afterwards and also it explicitly depends on the the implementation of the underlying classifier, while we want to be independent of it.

\cite{seo_neural_2018} use an LSTM with one full state vector and one reduced one. At each step they decide whether to use the reduced or the full state vector and can reduce computational cost by using the reduced one. For network traffic we think that it is often not even necessary to have a reduced state vector since from some attack flows, already the first couple of packets might contain enough information and then the remaining packets of the flow do not have to be considered anymore at all.

\cite{gui_long_2018} use \gls{rl} but their intention is not to skip samples but instead to make the \gls{rl} learn to choose a previous state that can aid at the current step. Thus it tries to help the \gls{rnn} with memorizing information from the past but does not actually take less data, which is the goal we pursue in this work.

These proposed techniques focus their evaluation on text and hence do not fulfill certain criteria we set out in \autoref{sec:introduction}. Furthermore, all techniques that use \gls{rl} use discrete actions, meaning that there is a hyperparameter $k$, which influences the maximum jump that is possible. However, for a network flow, already after the second packet it might be obvious that a flow is an attack and thus the remaining packets can be completely ignored. It is thus beneficial to not have a fixed maximum step size but a continuous one.

%\section{Setup}
%%ci

\section{Dataset}

\note{blue}{Fares: if we submit to CNS (page limit is 9), we can remove this sections as it is not relevant and statistics can be found online}

\begin{table}[h]
\caption{Occurrence frequency of the attack flows.}
\label{tab:occurrence}
\centering
\subfloat[CIC-IDS-2017\label{fig:cicids17proportions}
]{
\begin{tabular}{l r}
\toprule
Attack type & \hspace*{-4mm}Proportion \\ \midrule
DoS Hulk	&	10.10\%	\\
PortScan, Firewall	&	6.90\%	\\
DDoS LOIT	&	4.08\%	\\
Infiltration	&	3.30\%	\\
DoS GoldenEye	&	0.32\%	\\
DoS SlowHTTPTest	&	0.18\%	\\
DoS Slowloris	&	0.17\%	\\
Brute-force SSH	&	0.11\%	\\
Botnet ARES	&	0.03\%	\\
XSS attack	&	0.03\%	\\
PortScan, no Fw.	&	0.02\%	\\
Brute-force FTP	&	0.01\%	\\
SQL injection	&	$<$0.01\%	\\
Heartbleed	&	$<$0.01\%	\\
\bottomrule
\end{tabular}
}{}
\subfloat[UNSW-NB15\label{fig:unswnb15proportions}
]{
\begin{tabular}{l r}
\toprule
Attack type & \hspace*{-4mm}Proportion \\ \midrule
Exploits	&	1.42\%	\\
Fuzzers	&	1.01\%	\\
Reconnaissance	&	0.58\%	\\
Generic	&	0.21\%	\\
DoS	&	0.19\%	\\
Shellcode	&	0.08\%	\\
Analysis	&	0.03\%	\\
Backdoors	&	0.02\%	\\
Worms	&	0.01\%	\\
\bottomrule
\end{tabular}
}{}
\end{table}

For evaluation, we use the \textit{CIC-IDS-2017} \cite{sharafaldin_toward_2018} and \textit{UNSW-NB15} \cite{moustafa_unsw-nb15_2015} datasets, each includes more than 2 million flows of network data, containing both benign traffic and a large number of different attacks. The distribution of different attacks in the datasets are shown in \autoref{tab:occurrence}.

\section{\ours{}}

Traditionally, network packet sampling is performed on backbone infrastructures where, due to large bandwidths, not every packet can be analyzed. Sampling consequently helps experts understand the behavior of their network without the need to capture each and every packet. Assuming that the behavior of the network follows a predefined underlying process, sampling is therefore not detrimental on average to statistical analysis or measurement. Although the dataset in this work does not come from a backbone network, sampling reduces the high computational effort due to additional processing required by the complex \gls{ml}-based \glspl{ids}. As a consequence, several intuitive but efficient sampling techniques can be applied in practice. In this Section, we discuss the architecture of our proposed sampling framework \ours.

\subsection{Classifier}
\label{subsec:classifier}
We implemented a three-layer LSTM-based classifier with 128 neurons at each layer. As the input features we use source port, destination port, protocol identifier, packet length, \gls{iat} to the previous packet in the flow, packet direction (i.e. forward or reverse path) and all TCP flags (0 if the flow is not TCP).
We omitted \gls{ttl} values, as they are likely to lead to unwanted prediction behaviour \cite{bachl_walling_2019}.  Among the used features, source port, destination port and protocol identifier are constant over the whole flow while the others vary.
Additionally, we add the number of skipped packets since the last packet as a feature, which can help the classifier because otherwise it might be confused because of missing packets.
During flow extraction we used the usual 5-tuple flow key, which distinguishes flows based on the protocol they use and their source and destination port and IP address.
We use Z-score normalization and a train/test split of 2:1.

\subsection{Reinforcement Learning-based Sampling}
\label{subsec:rlForSampling}

Given enough training instances, \gls{rl} algorithms have been shown to perform very well in complex decision-making situations such as playing video games \cite{mnih_playing_2013} as well as networking problems such as congestion control \cite{bachl_rax_2019}. We thus consider an \gls{rl}-based architecture to be suitable for taking packet sampling decisions. The system should look at the current packet as well as the history of the conversation flow and output how many packets should be skipped until the next packet is considered in the current flow. The reward that the \gls{rl} tries to maximize should be a combination of classification accuracy and the number of packets that could be skipped (sparsity). An operator should be able to specify how much accuracy is allowed to be traded off for achieving higher sparsity.% and the objective of the \gls{rl} at a timestep t of a flow is defined as:

%\begin{align}
%\begin{split}
%U_t = & \text{future average classification performance} \\
%+ & \alpha\times\text{future average sparsity}
%\end{split}
%\end{align}

For a flow of length $T$ at each timestep $t$ we have two reward metrics which we compute in a fashion that is known from \textit{R-learning} \cite{schwartz_reinforcement_1993}:

%\begin{figure}
%\centering
%  \includegraphics[width=\columnwidth]{img/rewards.pdf}
%  \caption{The computation of the rewards for a sample flow: Gray packets are skipped while white ones are chosen. At the top row the resulting rewards for classification for each chosen packet are depicted and in the bottom row the ones for sparsity. Numbers are rounded to two digits after the decimal point.}
%  \label{fig:rewards}
%\end{figure}

\begin{figure*}
\centering
  \includegraphics[width=0.8\textwidth]{img/rewards_calc_example.pdf}
  \caption{The computation of the rewards for a sample malicious flow: Gray packets (0) are skipped while black ones (1) are chosen. ``\#Left\//\#Total'' and ``Confidence'' are the underlying parameters for rewards calculations (see \autoref{eq:R_classification} and \autoref{eq:R_sparsity}). Numbers are rounded to two digits after the decimal point.}
  \label{fig:rewards}
\end{figure*}

The classification reward metric (\autoref{eq:R_classification}) captures how correct the classifier is on average for all future packets of a flow. The label of each packet in a flow has a value of 0 for non-attack traffic and 1 for attack traffic with one flow being only attack or non-attack but not mixed. The confidence is the sigmoided $\left(\frac{e^x}{1+e^x}\right)$ output of the classifier, yielding a value between 0 and 1, where 0 is absolute confidence for a flow being benign and 1 is absolute confidence for a flow being an attack.
\begin{align}
\begin{split}
R_{\text{classification},t} = \frac{1}{T-t} \sum_{i=t+1}^{T} |1 - \left(\text{label}_i - \text{confidence}_i\right)|
\end{split}
\label{eq:R_classification}
\end{align}

The sparsity reward metric (\autoref{eq:R_sparsity}) captures how much sparsity is achieved on average for all future packets of a flow. E.g. for a flow of length 10 if after timestep 5, 3 packets are chosen and 2 are skipped from what is left, the sparsity reward metric at timestep 5 is $\frac{2}{10-(4+1)}=\frac{2}{5}$.
%\note{blue}{Fares: This equation can be shortened by diving by the total count of packets left.\\
%Fares: If the length of the flow is not known, how do we estimate T}.
%\note{red}{Max: It's only relevant during training when you know the lengths of the flows. In deployment, you don't need the rewards anymore...}
\begin{align}
\begin{split}
R_{\text{sparsity},t} = \frac{\sum_{i=t+1}^{T} 1_{\text{skipped},i}}{T-(t+1)}
\end{split}
\label{eq:R_sparsity}
\end{align}

\autoref{fig:rewards} shows how classification and sparsity rewards are computed for a toy example.

We opt for an \emph{\gls{ac}} approach (as shown in \autoref{fig:sampling}) modeled after \cite{mnih_asynchronous_2016} because we consider it being more interpretable than a Deep \gls{rl} approach based on a \textit{Deep Q Network} like \cite{mnih_playing_2013} for reasons we explain later \todo{Max: Add ref to the section where we explain this}.

Our proposed framework (see \autoref{fig:neuralNetworkArchitecture}) consists of three independent neural networks: the classifier, the critic and the actor.

The classifier's goal is to guess correctly if a flow is an attack or not. For this, it outputs its confidence for non-attack/attack after reading a packet. As stated in \autoref{subsec:classifier}, we choose an LSTM-based supervised classifier for this work, but our framework can work together with any classifier as long as it outputs a confidence for non-attack/attack at each timestep and can handle skipped packets. For instance it would also be possible to replace the LSTM with an unsupervised autoencoder like in \cite{mirsky_kitsune_2018}.

The Critic aims to estimate the future expected average classification performance and the future expected average sparsity that can be achieved at a timestep $t$. For this it outputs $v_{\text{classification},t}$ and $v_{\text{sparsity},t}$ at each timestep $t$. It aims to predict the classification performance and sparsity correctly for each timestep by minimizing the following value loss function:
\begin{align}
\begin{split}
l_{\text{v},t} =& \left(R_{\text{classification},t} - v_{\text{classification},t}\right)^2 \\
+ & \left(R_{\text{sparsity},t} - v_{\text{sparsity},t}\right)^2
\end{split}
\end{align}

Now that the reward metrics and the Critic's outputs are defined, we can define the overall utility $U$ that we want to maximize at each timestep $t$ as follows:
\begin{align}
\begin{split}
U_t =& \left(R_{\text{classification},t} + \alpha \cdot R_{\text{sparsity},t}\right) \\
- & \left(v_{\text{classification},t} + \alpha \cdot v_{\text{sparsity},t}\right)
\end{split}
\end{align}

This means that we want the reward to be higher than the Critic's expectation.

The Actor outputs a probability distribution at timestep $t$, which is sampled to determine the number of packets that should be skipped. It aims to change the distribution so that actions which result in a higher reward are chosen more frequently, maximizing the utility function. However, this strategy could lead to the Actor getting stuck with what it considers the best decision and never try alternatives even though they might result in a higher reward (local maxima). Thus, the Actor not only tries to change the distribution so that the actions become more optimal (higher reward) but at the same time also aims to maximize the entropy of the distribution so that alternative choices are still explored reasonably often.

Specifically, at timestep $t$, given the current state $s_t$ (which consists of the current input as well as the LSTM state) and the neural network weights of the Actor $\theta_a$ the Actor outputs a probability distribution $\pi \left( a_t \given s_t ; \theta_\text{a} \right)$ from which the action $a_t \ge 1$ is sampled. $a_t=1$ means that the next packet is inspected next, while, for example, $a_t=2$ means that one packet is skipped until the next that is chosen and so forth. Besides wanting to make actions which result in a high reward more likely, the actor also seeks to optimize the entropy $H(\cdot)$ of the probability distribution so that it keeps exploring and doesn't get stuck with a suboptimal policy.

Combining these objectives, the actor network aims to minimize the loss function
\begin{align}
\begin{split}
l_{\text{a},t} =& \underbrace{ -\log \left( \pi \left( a_t \given s_t ; \theta_\text{a} \right) \right) U_t}_{\text{Policy Loss}} \\
&- \beta \underbrace{H\left( \pi\left( s_t; \theta_a \right)\right) }_{\text{Enthropy}}
\end{split}\label{eq:actor}
\end{align}

%The Critic network evaluates the behavior of the Actor network and computes the classification performance and sparsity that it thinks is going to be achieved.% using the computed value function and gives it feedback based on the performance obtained by the LSTM classifier.
%The number of packets to skip is taken from the Actor's output distribution and fed into the packet parser, which is responsible for counting the packets and selecting the appropriate one.

In related work regarding \gls{rl}-based sampling for sequences, only discrete actions were evaluated so far: The actor would output a categorical probability distribution with a fixed maximum number of bins $k$. A drawback of this approach is that only a predefined maximum number of packets can be skipped and that for a very large number of $k$, it can take a long time to converge to an optimum policy as there are $k$ different options to explore at each timestep $t$. Besides discrete actions, we thus also experiment with continuous actions, which allow to skip arbitrarily many packets.

\begin{figure}
\centering
  \includegraphics[width=\columnwidth]{neural_network.pdf}
  \caption{An overview of the complete neural network being used in our experiments. Ocher stands for the input, blue for linear layers, red for LSTM layers and purple for the outputs of the neural network. The numbers in parentheses next to each label stand for the width of that layer (i.e.~the number of neurons (except for the inputs and outputs)). Parts of the neural network are shared between $v_\text{classification}$ and $v_\text{sparsity}$ as well as $\mu$ and $\sigma$ as they output quantities of similar magnitude and thus the gradients that are computed when updating the neural network are of the same order of magnitude as well. Our implementation also supports sharing the LSTM between all three neural network. However, as the five outputs can produce losses that have different order of magnitude, larger gradients or larger losses would ``overwrite'' the gradients of the smaller ones.}
  \label{fig:neuralNetworkArchitecture}
\end{figure}

%\note{blue}{Fares: can be moved one section above\\}
%An overview of the neural networks of the classifier, actor and critic that we use in this work is given in \autoref{fig:neuralNetworkArchitecture}.

\begin{figure}
\centering
  \includegraphics[width=0.8\columnwidth]{img/rnn-sampling.pdf}
  \caption{SparseIDS framework. Three deep networks are shown: an LSTM classifier, An Actor network and a Critic network. The Actor takes actions based on the feedback from the the Critic network and the classifier. The Critic network tracks the Actor actions based on the classifier confidence and finally, the packets selector parses packets based on the decision of the Actor network.}
  \label{fig:sampling}
\end{figure}

%\note{blue}{Fares: can be removed, it was already mentioned.\\}
%A schematic overview over the \gls{rl} for sampling is given in \autoref{fig:sampling}.

\subsection{Continuous Actions}

For performing continious actions we need a probability distribution which can be parametrized to have the mean between 0 and positive infinity. Furthermore it must have at least two parameters so that the standard deviation, which is required for the entropy (\autoref{eq:actor}), can be changed independent of the mean. For example, the Exponential distribution can be parametrized to have its mean anywhere between 0 and positive infinity, however, it has only one parameter and thus the mean and the standard deviation cannot be independently changed. Conversely, a simple distribution which fits our requirements is the Log-normal distribution, which is defined as a normal distribution whose output is exponentiated:

Let $Z$ be a random variable from a standard normal distribution $\mathcal{N}(0, 1)$ and let $\mu_{\text{normal}}$ and $\sigma_{\text{normal}} > 0$ be two real numbers. Then
\begin{align}
X = e^{\mu_{\text{normal}}+\sigma_{\text{normal}} Z}
\end{align}
is log-normally distributed variable from $\mathcal{LN}(\mu, \sigma^2)$.

The architecture in \autoref{fig:neuralNetworkArchitecture} shows that the actor network has two outputs: $\mu$ and $\sigma$. These encode the mean and standard deviation of a log-normal distribution and can never be smaller or equal to zero hence, we always apply a rectifier function at the output. We choose the softplus, which is a smooth rectifier function:
\begin{align}
\text{softplus}(x) = \log\left(1 + e^x \right)
\end{align}
To parametrize the normal distribution that underlies the log-normal distribution, we first have to compute the parameters $\mu_{\text{normal}}$ and $\sigma_{\text{normal}}$ from $\mu$ and $\sigma$. We can compute them as follows:

\begin{align}
\mu_{\text{normal}} = \frac{1}{2}\left( 2\log(\mu) - \log\left(\frac{\mu^2 + \sigma^2}{\mu^2}\right) \right)
\end{align}

\begin{align}
\sigma_{\text{normal}} = \sqrt{\log\left(\frac{\mu^2 + \sigma^2}{\mu^2}\right)}
\end{align}

When experimenting with continuous actions, we encountered one practical problem: At the beginning of training, the classifier would still perform badly and would not give the actor a useful reward. As a result, the actor would only focus on and maximize sparsity, as the classification reward would always be equally bad no matter how many packets are chosen. Thus, the actor would learn to let $\mu$ go to infinity as this would increase the reward at the beginning when the interaction between the three components of our \gls{rl} system is still fragile. As a solution, we add a penalty to the reward calculation when the actor jumps far behind the last flow packet.The modified reward applies only to the last chosen packet and penalizes the actor jumping more than one packet behind the last packet. The sparsity reward formula (\autoref{eq:R_sparsity}) of the last packet is modified as follows, where \textit{last} refers to the index of the last chosen packet in the flow:

\begin{align}
\begin{split}
R_{\text{sparsity},\textit{last}} = \frac{\left(\sum_{i=\textit{last}+1}^{T} 1_{\text{skipped},i}\right) - \left(\textit{last}+a_{\textit{last}} - T - 1\right)}{T- (\textit{last}+1)}
\end{split}
\label{eq:R_sparsity_last}
\end{align}

Specifically, it adds the term $\left(\textit{last}+a_{\textit{last}} - T - 1\right)$ to the numerator which penalizes the actor jumping more than one packet behind the flow. Jumping just right after the flow does not get penalized because if the actor considers the rest of the flow as irrelevant, ignoring the rest of the flow and jumping to the first packet behind the flow is a useful action.

\section{Other sampling techniques}
\label{sec:other_sampling_techniques}
In order to evaluate the performance of \ours, we compare it with a few rule-based sampling techniques widely used in practice.
We opt for three sampling families that do not require feedback from higher layers (i.e. sampling rates are constant and not correlated with subsequent analysis): %The underlying principle of each family is summarized as follows:

\subsection{Random Sampling}

Given the fraction $p$ of packets to be selected in a sequence of $T$ packets, this technique will sample each individual packet with probability of $p$. If the sequence is large enough, $n$ packets will be randomly sampled from a uniform distribution whereby $n= p \times T$. Practically, these techniques work properly (the effect on subsequent analysis is minimal) if the candidate packets have similar properties (redundancy) and therefore the probability of skipping important information is minimal.

\subsection{First \emph{n} Sampling}
As its name suggests, this technique only samples the first packets from each flow and ignores the rest of the flow. For a sampling probability $p$ it will take the first $n$ packets from each flow so that $n \approx T * p$. For example, for a flow with length 10 and $p=0.2$, it will take the first 2 packets. For this procedure the length of a flow has to be known beforehand and thus it is not applicable in a streaming scenario. This technique is successful under the assumption that most information for determining maliciousness is present in the first few packets of a flow, i.e. a large data transfer in which the initial packets represent a hand-shake after which the majority of packets contain the raw data and do not reveal significant information to non-\gls{dpi} systems.

\subsection{First \emph{n} Equal Sampling}
This technique is similar to the one above, with the difference that the length of a flow doesn't have to be known beforehand. If flows have an average length of 10 packets, and the sampling probability $p=0.2$ it would take up to two packets from each flow.

\subsection{Every \emph{i}th Sampling}
Finally we use a periodic sampling technique that takes a sample each $i= 1/ \alpha$ packet. Similar to the last technique, the length of the flow is not needed in this case. In practice, this technique is best suited to keep the behavioral distribution of the sampled set similar to the original distribution, thus suited for scenarios where information is heterogeneously distributed between packets over an entire network flow.

In each of the above techniques, the first packet of each flow is always taken because otherwise one could get undefined behavior: If a flow only consists of 1 packet and this packet is not chosen, it is not even defined if the prediction of the classifier was correct for this flow or not. Thus, in our implementations of the above sampling techniques, we make sure that even when the first packet is always chosen, the overall fraction of chosen packets corresponds to the sampling probability $p$.

After applying the sampling techniques, the sequence of selected packets is then fed to the classifier discussed in \autoref{subsec:classifier}, where features are extracted and a decision is made after each packet. Also, the classifier gets information on how many packets were skipped prior to each packet.

% Max: If we write things like this they might think we were too lazy to implement them :D
%There exist many more rule-based sampling techniques e.g., temporal, adaptive etc. However, studying their behaviors is out of the scope of this paper.

\section{Results}

The results of the sampling part are discussed in this section. We study the sparsity-accuracy tradeoff first, then we evaluate both discrete and continuous actions and, finally, we compare our results with other sampling techniques. Overall, sampling based on \gls{rl} reveals that a high number of packets can be skipped while maintaining high performance.

\begin{table}
% Evaluated using
% ./python learn.py --dataroot flows.pickle --net runs/Oct26_00-03-50_gpu/lstm_module_1284.pth --function test --batchSize 512
% ./python learn.py --dataroot flows15.pickle --net runs/Oct28_15-41-46_gpu/lstm_module_997.pth --function test --batchSize 512
% on Dec. 4th
\caption{Comparing classification metrics per flow with different tradeoffs for both CIC-IDS-2017 and UNSW-NB15.} \label{tab:results_tradeoff}
\centering
\begin{tabular*}{\columnwidth}{>{\rowmac}l @{\extracolsep{\fill}} >{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c<{\clearrow}} \toprule
& \multicolumn{3}{c}{CIC-IDS-2017} & \multicolumn{3}{c}{UNSW-NB15} \\
Tradeoff & 0.0 & 0.1 & 0.5 & 0.0 & 0.1 & 0.5 \\	\midrule
\setrow{\bfseries}Sparsity & 0\% & 76.3\% & 76.5\% & 0\% & 83.5\% & 84.5\% \\ \midrule
\setrow{\bfseries}Accuracy & 99.5\% & 99.4\% & 99.3\% & 98.9\% & 98.4\% & 98.5\% \\
Precision & 99.4\% & 98.5\% & 98.7\% & 81.1\% & 69.5\% & 70.8\% \\
Recall & 98.5\% & 99.0\% & 98.6\% & 84.0\% & 89.9\% & 89.6\% \\
F1 & 99.0\% & 98.7\% & 98.6\% & 82.5\% & 78.4\% & 79.1\% \\
Youden & 98.4\% & 98.5\% & 98.2\% & 83.3\% & 88.6\% & 88.4\% \\ 
\bottomrule
\end{tabular*}
\end{table}

\subsection{Influence of the tradeoff}

\autoref{tab:results_tradeoff} shows the influence of the tradeoff parameter $\alpha$ for both datasets. For columns with a tradeoff of 0.0, we train the classifier regularly with all packets without using \gls{rl}. For tradeoffs 0.1 and 0.5 we train \ours{} using \gls{rl} with continuous actions and keep the tradeoff constant for the whole training. We train for 8 epochs ($>10$ million flows for both datasets) and then evaluate the resulting neural networks on the test dataset.\par
Results show that the higher the tradeoff, the less packets are sampled and the lower the accuracy. The only exception to this is that for tradeoffs of $0.1$ and $0.5$ for UNSW-NB15, the accuracy is almost the same or even slightly higher for the larger tradeoff, which is counter-intuitive. We consider this to be a coincidence that can happen because of the stochasticity of the training process.

\note{blue}{Fares: my suggestion: 
The only exception to this is that for tradeoffs of $0.1$ and $0.5$ for UNSW-NB15, the accuracy is almost the same or even slightly higher for the larger tradeoff, which is counter-intuitive. We assume that due to the nature of the UNSW-NB15 data set (which is highly heterogeneous), fewer packets yield better information (and therefore the memory of the LSTM classifier is cleaner) and hence better classification decision and higher accuracy. 
}

Furthermore, due to the fact that UNSW-NB15 contains more packets per flow, its sparsity is higher. This, on the other hand, holds since almost the same number of packets is sampled from each flow for both datasets (see \autoref{sec:conclusion}).

%\todo{Write that the sparsity is higher for UNSW-NB15 because it contains longer flows}

\subsection{Discrete vs.~Continuous actions}

\begin{table}
\caption{Comparing classification metrics per flow with either discrete (20 different actions) or continuous actions for both CIC-IDS-2017 and UNSW-NB15. The tradeoff is always $0.1$.} \label{tab:results_disc_cont}
\centering
\begin{tabular*}{\columnwidth}{>{\rowmac}l @{\extracolsep{\fill}} >{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c<{\clearrow}} \toprule
& \multicolumn{2}{c}{CIC-IDS-2017} & \multicolumn{2}{c}{UNSW-NB15} \\
Actions & cont. & disc. & cont. & disc. \\	\midrule
\setrow{\bfseries}Sparsity & 76.3\% & 73.6\% & 83.5\% & 86.6\% \\ \midrule
\setrow{\bfseries}Accuracy & 99.4\% & 99.2\% & 98.4\% & 98.4\% \\
Precision & 98.5\% & 98.7\% & 69.5\% & 68.0\% \\
Recall & 99.0\% & 98.3\% & 89.9\% & 94.6\% \\
F1 & 99.7\% & 98.5\% & 78.4\% & 79.2\% \\
Youden & 98.5\% & 97.8\% & 88.6\% & 93.2\% \\ 
\bottomrule
\end{tabular*}
\end{table}

\autoref{tab:results_disc_cont} compares continuous and discrete (20 different) actions and shows that for CIC-IDS-2017 continuous actions result in a higher accuracy with fewer samples being selected. Conversely, for UNSW-NB-15, accuracy is the same for both continuous and discrete actions. However, with discrete actions, fewer packets are chosen, which leads to the conclusion that in this case, discrete actions perform better. This leads to the conclusion, that at least for our use case and the two datasets that we consider, there is no consistent difference in performance for continuous and discrete actions. Generally, the advantage of discrete actions is that they are theoretically more powerful as they can compare each of the possible actions with each other, while the only possibility for continuous actions is to increase or decrease the mean log-normal distribution, which can lead to worse performance. On the other hand, with discrete actions, the maximum number of samples to skip is fixed, which we consider to be a major drawback. Furthermore, training time is potentially higher since, for example, for a discrete space of 20 actions, each option is treated as being completely different while actually there is not a large difference between skipping 18 or 19 packets, there is a big difference between skipping 0 packets compared to skipping 19 packets. This means that for discrete actions, all actions are perceived as being completely independent, which is not true as the reward function considers sparsity, which increases gradually as more packets are skipped and for which similar actions result in similar rewards.

\subsection{Comparing with other sampling methods}

\begin{table}
\caption{Comparing classification metrics per flow between \gls{rl}-based sampling with continuous actions (tradeoff $0.1$) and alternative strategies for CIC-IDS-2017.} \label{tab:results_others}
\centering
\begin{tabular*}{\columnwidth}{>{\rowmac}l @{\extracolsep{\fill}} >{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c<{\clearrow}} \toprule
& \gls{rl} & random & every $i$th & first $n$ & first $n$ equal \\	\midrule
\setrow{\bfseries}Sampled & 76.3\% & 76.3\% & 76.3\% & 76.3\% & 76.3\% \\ \midrule
\setrow{\bfseries}Accuracy & 99.4\% & 96.9\% & 97.8\% & 97.3\% & 98.3\% \\
Precision & 98.5\% & 92.4\% & 95.8\% & 93.8\% & 95.6\% \\
Recall & 99.0\% & 95.6\% & 95.4\% & 95.5\% & 97.9\% \\
F1 & 99.7\% & 94.0\% & 95.6\% & 94.7\% & 96.7\% \\
Youden & 98.5\% & 93.0\% & 93.9\% & 93.4\% & 96.4\% \\
\bottomrule
\end{tabular*}
\end{table}

Next we compared the \gls{rl} technique usedy by \ours{} with several intuitive sampling strategies suggested in \autoref{sec:other_sampling_techniques}. 
\autoref{tab:results_others} reveals that \ours{} achieves the highest accuracy with the same sparsity. Not surprisingly, random sampling performs the worst, as it is likely that a flow is either ``misfortune'' and doesn't get enough packets sampled while another  one gets more or, the random packets chosen carry less useful information. The best strategy besides the \gls{rl} approach seems to be ``first $n$ equal'' sampling, which matches the conclusion from \cite{hartl_explainability_2019} that the first couple of packets are the most important ones for determining maliciousness of a flow. 
\note{blue}{Fares: The next sentence is a bit confusing.}
``first $n$ equal'' performs better than ``first $n$'' because for ``first $n$'', longer flows will have more packets sampled from them. Conversely, very short flows will get few packets. But since the first packets of a flow determine its maliciousness, the short flows get misclassified because they don't get enough packets sampled.

\section{Steering}

So far, we have always trained \ours{} with a fixed tradeoff parameter $\alpha$. Nevertheless, there is also an important use case in which an operator wants a certain minimum sparsity, for example 50\%, to limit the necessary computation power and thus reduce costs. 
We thus propose a method that enables steering the tradeoff to achieve a certain desired sparsity.
\begin{enumerate}
\item A model is trained where the tradeoff is randomly chosen for each flow during training. We set the probability distribution to be a uniform distribution between 0 and 1. We add a new feature to each packet which indicates the tradeoff chosen for this flow so that it can adapt its sampling strategy in anticipation of the tradeoff that will determine the reward.
\item During deployment, the tradeoff is set to the maximum that was used during training (1 in our case), which results in maximum sparsity. Then, the tradeoff is continuously incremented if the current sparsity is still under the specified minimum. It is important that the tradeoff can never fall below 0 since this is a case that was not encountered during training and which can lead to anomalous behavior. 
\end{enumerate}

%{{"../plots/plot/flows_pred_plots2_outcomes_0_3.pickle_5_DoS - DDoS-DoS Hulk"}
\begin{figure}
\centering
  \includegraphics[width=\columnwidth]{{"img/lstm_module_8.pth_global_tradeoff_1.0_steering_step_size_0.05_steering_target_sparsity_0.5_batches_to_consider_for_steering_500_sampling_rl"}.pdf}
  \caption{Steering the tradeoff with an allowed minimum sparsity of 50\%. The tradeoff is initialized as 1 in the beginning and continuously lowered as long as the sparsity is higher than 50\% for CIC-IDS-2017. The procedure stops if the minimum sparsity is reached or the tradeoff reached its minimum of 0, which is the case here. Sparsity falls from around 74.5\% to 73.5\%.}
  \label{fig:steering}
\end{figure}

\autoref{fig:steering} shows the steering in action. It stops when the tradeoff reaches 0 and the specified minimum sparsity of 50\% has not been reached yet. It might seem surprising that even with a parameter of 0, which means that only classification matters and sparsity not at all, still a sparsity of 73.5\% is achieved. We attribute this to the fact that \ours{} learns that some flows contain may contain packets that do not contain any useful information about determining maliciousness. Thus they are even skipped when there is no reward for skipping them because they do not increase the classification score at all.

\section{Conclusion}
\label{sec:conclusion}

We saw that the \gls{rl}-based approach with a tradeoff between classification accuracy and sparsity works as expected and that higher tradeoffs resulted in higher sparsity and potentially lower classification performance. Surprisingly, we saw that our solution \ours{} chooses very few packets: Typically $\approx \frac{1}{4}$ for CIC-IDS-2017 and $\approx \frac{1}{5}$ for UNSW-NB-15.

Furthermore, our results showed that \gls{rl}-based sampling can intelligently choose relevant packets for intrusion detection and outperform non-smart sampling techniques. 
Besides discrete actions for our \gls{rl} system we also implemented continuous ones. A naive implementation did not work as the training was too fragile in the beginning and the \gls{rl} system would diverge and the number of skipped packets would go to infinity. Adding a penalty for jumping behind the end of the sequence proved effective. Comparing with discrete actions, continuous actions seemed to perform on par. This is still an advantage, as for continuous actions the neural network only needs to have two outputs (mean and standard deviation) while for discrete ones it needs many outputs as there are actions. The continuous action framework is thus more lightweight.

\section*{Acknowledgements}
The Titan Xp used for this research was donated by the NVIDIA Corporation.

\renewcommand*{\bibfont}{\small}
\bibliographystyle{ieeetr}
\bibliography{bibliography}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.