### 2020-02-06 paper_rl/paper.tex
Remove UNSW-NB15 for now and bring it under 9 pages

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[conference]{IEEEtran}
%% IEEE CNS addition:
\makeatletter
\def\ps@headings{%
\def\@oddhead{\mbox{}\scriptsize\rightmark \hfil \thepage}%
\def\@evenhead{\scriptsize\thepage \hfil \leftmark\mbox{}}%
\def\@oddfoot{}%
\def\@evenfoot{}}
\makeatother
\pagestyle{empty}

\usepackage[utf8]{inputenc}
\usepackage{multirow}
\usepackage[inline]{enumitem}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{subfig}

\usepackage[nomain, toc, acronym]{glossaries}
\glsdisablehyper

\newcommand\note[2]{{\color{#1}#2}}
\newcommand\todo[1]{{\note{red}{TODO: #1}}}

% Adds a "given" symbol (vertical bar) that rescales in height (like \left and \right)
\usepackage{mathtools}
\newcommand\givenbase[1][]{\:#1\lvert\:}
\let\given\givenbase
\newcommand\sgiven{\givenbase[\delimsize]}
\DeclarePairedDelimiterX\Basics[1](){\let\given\sgiven #1}
\newcommand\Average{E\Basics}

\usepackage{tabularx}

% Macro to make entire row bold from https://tex.stackexchange.com/questions/309833/format-whole-row-of-table-as-bold
\newcommand\setrow[1]{\gdef\rowmac{#1}#1\ignorespaces}
\newcommand\clearrow{\global\let\rowmac\relax}
\clearrow

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{SparseIDS: Learning Packet Sampling with Reinforcement Learning}

\author{\IEEEauthorblockN{Maximilian Bachl, Fares Meghdouri, Joachim Fabini, Tanja Zseby}
\IEEEauthorblockA{Technische Universität Wien\\
firstname.lastname@tuwien.ac.at}}

%\date{\today}

% \IEEEoverridecommandlockouts
% \IEEEpubid{\begin{minipage}[t]{\textwidth}\ \\[10pt]
%         \centering\normalsize{xxx-x-xxxx-xxxx-x/xx/\$31.00 \copyright 2018 IEEE}
% \end{minipage}}

% \renewcommand*{\bibfont}{\footnotesize}

\maketitle%

%\thispagestyle{plain}
%\pagestyle{plain}

\newacronym{ml}{ML}{Machine Learning}
\newacronym{dl}{DL}{Deep Learning}
\newacronym{aml}{AML}{Adversarial Machine Learning}
\newacronym{ids}{IDS}{Intrusion Detection System}
\newacronym{rnn}{RNN}{Recurrent Neural Network}
\newacronym{fgsm}{FGSM}{Fast Gradient Sign Method}
\newacronym{cw}{CW}{Carlini-Wagner method}
\newacronym{pgd}{PGD}{Projected Gradient Descent}
\newacronym{pdp}{PDP}{Partial Dependence Plot}
\newacronym{ars}{ARS}{Adversarial Robustness Score}
\newacronym{ttl}{TTL}{Time-to-Live}
\newacronym{dos}{DoS}{Denial-of-Service}
\newacronym{iat}{IAT}{Interarrival time}
\newacronym{rl}{RL}{Reinforcement Learning}
\newacronym{dpi}{DPI}{Deep Packet Inspection}
\newacronym{ac}{AC}{Actor-Critic}

\newcommand{\ours}{SparseIDS}

\begin{abstract}

\glspl{rnn} have been shown to be valuable for constructing \glspl{ids} for network data. They already allow determining if a flow is malicious or not before it is over, making it possible to take action immediately. However, considering the large number of packets that have to be inspected, the question of computational efficiency arises. We show that by using a novel \gls{rl}-based approach called \textit{\ours{}}, we can reduce the number of consumed packets by more than three fourths while keeping classification accuracy high. Comparing to various other sampling techniques, \ours{} consistently achieves higher classification accuracy with the same number of sampled packets. A major novelty of our \gls{rl}-based approach is that it can not only skip up to a predefined maximum number of samples like other approaches proposed in the domain of Natural Language Processing but can possibly skip arbitrarily many packets in one step, saving even more computational resources for long sequences. Inspecting \ours{}'s behavior of choosing packets shows that it adopts different sampling strategies for different attack types and network flows. Finally we build an automatic steering mechanism that can guide \ours{} in deployment to achieve a desired level of sparsity.

\end{abstract}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
%\begin{abstract}
%  A clear and well-documented \LaTeX\ document is presented as an
%  article formatted for publication by ACM in a conference proceedings
%  or journal publication. Based on the ``acmart'' document class, this
%  article presents and explains many of the common variations, as well
%  as many of the formatting elements an author may use in the
%  preparation of the documentation of their work.
%\end{abstract}

\maketitle

\section{Introduction}
\label{sec:introduction}
\note{blue}{Fares: I fixed the paper based on Tanja's comments. still  (1) the abstract is not touched yet (2) figures with axes (time step) need to be changed... We need to be carefull with the 9 pages limit}

%There is a significant body of scientific work focusing on the detection of unwanted behavior in networks. In the past, a viable way of performing intrusion detection was to inspect the content of packets themselves and detect if a packet delivers potentially harmful content. More recently, with the increasing deployment of encryption, the focus now lies on features that are always available to network monitoring equipment like port numbers, protocol flags or packet sizes when encrypting above the transport layer.

% Fares:
As the need for security mechanisms in today's infrastructure's communication networks grows, a substantial body of scientific work focuses on developing reliable and explainable frameworks for the identification of unwanted behavior. In the past, a viable way to detect intrusions was to analyze the contents of the packets themselves and determine, for example, whether a packet contains potentially harmful content by matching patterns. More recently, with the increasing deployment of encryption, the focus is now on features available for network monitoring devices even if packets are encrypted, such as packet sizes and temporal behavior. In our work we concentrate on features available when encrypting above transport layer (like for example TLS or QUIC), i.e. we also include port numbers and protocol flags.

%Network communication is usually aggregated into \textit{flows}, which are commonly defined as a sequence of packets sharing certain properties. When analyzing flows, not only the aforementioned features are available but also features related to the timing of the individual packets. Various approaches have been proposed to extract features from flows and then perform anomaly detection with the extracted flows \cite{meghdouri_analysis_2018}.
%While these approaches frequently work well, it is problematic that the whole flow has to be received first and only afterwards anomaly detection is applied, revealing attack flows. Thus, we design a network \gls{ids} that operates on a per-packet basis and decides if a packet is anomalous based on features that are available even for traffic that is encrypted above the transport layer, like for example TLS or QUIC.
%At the same time, an \gls{rnn}-based \gls{ids} has the benefit of avoiding tedious feature engineering procedures, which derive statistical measures from the sequence of packet features, but providing any available information to the classifier.
%\cite{hartl_explainability_2019} showed that such a system has similar performance to other flow-based anomaly detection systems but can detect anomalies \textit{before} the flow terminates and analyzed its security with respect to adversaries extensively.

% Fares:
Network communication is typically aggregated into \textit{flows}, which are commonly defined as a sequence of packets that share certain properties. When analyzing flows, not only the aforementioned features are available but also features representing the temporal behavior of the individual packets. Various approaches have been proposed to extract flow features with which anomaly detection can be performed: Authors in \cite{meghdouri_analysis_2018} compare such approaches and show their efficacy.
While these approaches often work well, a major drawback is that the whole flow must be received first and only then anomaly detection can be applied, to reveal malicious flows. Thus, we design a network \gls{ids} that operates on a per-packet basis and determines whether a packet is anomalous based on the aforementioned features.
In this way, an \gls{rnn}-based \gls{ids} has the benefit of avoiding feature engineering procedures and let the classifier build high level features by itself.
\cite{hartl_explainability_2019} show that such an architecture has similar performance to traditional flow-based anomaly detection systems, but can still detect anomalies \textit{before} the flow ends. Furthermore they analyzed its robustness with respect to adversaries extensively.

%However, for practical use, not only classification accuracy and security are important but also computational efficiency. This follows from the fact that a rationally-acting company will only deploy an \gls{ids} if the cost that is saved by avoiding security breaches is higher than the cost of the \gls{ids} itself. The cost of the \gls{ids} can be split in the cost of purchase/installation, the cost of operating it and the cost of maintaining it. The goal we pursue here is to develop sampling strategies that do not significantly degrade classification performance while being able to only process a small fraction of the original data. Specifically, the sampling technique should

% Fares:
Nevertheless, not only the accuracy and reliability of classification, but also the computational efficiency, are critical for practical use. This follows from the fact that a rationally-acting organization will only implement an \gls{ids} if the cost saved by preventing security breaches is higher than the cost of the \gls{ids} itself. The cost of an \gls{ids} can be divided into (a) the cost of purchase/installation, (b) the cost of operation and (c) the cost of maintenance. The goal we pursue here is to develop optimal sampling strategies that do not significantly degrade classification performance while being able to only process a small fraction of the original data. Specifically, a good sampling technique should:

\begin{enumerate}
%\item \textbf{choose optimally}, taking only the data that contain the most information and skip the less relevant ones.
% Fares:
\item \textbf{choose optimally}, taking only samples that contain the most information and skip the less relevant ones.
\item have a parameter that allows to \textbf{trade off} classification performance for sparsity (choosing fewer packets).
\item be \textbf{independent of the classifier} so that any classifier can be used and that the classifier doesn't have to be aware of the sampling strategy.
\item be \textbf{retrainable} so that the sampling can be continuously adapted depending on the current threat landscape.
\item be able to \textbf{skip arbitrarily} many packets since network flows can be very long but only the first couple of packets are needed to decide whether an attack occurred or not and the remaining packets can be skipped altogether.
\end{enumerate}

%For this purpose, we develop the \gls{rl}-based \gls{ids} \textit{\ours{}} fulfills the above properties. We show that a significant number of packets can be skipped while accuracy doesn't drop significantly. Compared to other common sampling strategies, \ours{} performs better when being trained on the same number of packets.

% Fares:
For this purpose we develop an \gls{rl}-based \gls{ids}: \textit{\ours{}} that fulfills the above properties. We show that a significant number of packets can be skipped while the accuracy does not drop
considerably. Compared to other common sampling techniques, \ours{} performs better when being trained on the same number of packets.

%As \ours{} uses a parameter to trade-off accuracy for sparsity, the resulting accuracy and sparsity are determined by this parameter. However, since it also might be desired to specify the sparsity itself, we propose a steering system which takes a desired sparsity as its input and modifies the parameter of the reinforcement learning in a control loop until the specified sparsity is reached.

% Fares:
The accuracy-sparsity trade-off of \ours{} can be regulated by a reinforcement learning parameter that directly affects both quantities. However, as in some deployments, a certain sparsity has to be minimally achieved to limit the computational expense, we also develop a steering system that adapts the trade-off parameter in a closed control loop. In this way the given budget of computation power is not exceeded.

To encourage reproducibility and facilitate experimentation, we publicly release the source code, the trained ML models, the data and the figures of this work\footnote{\url{https://github.com/CN-TU/adversarial-recurrent-ids/tree/rl}}.

\section{Related Work}

There is a significant body of research on sampling for networks: \cite{zseby_sampling_2005} give an overview of common methods for packet sampling in IP networks. \cite{estan_building_2004} develop a scheme that adapts the sampling rate depending on available resources.

Several approaches have been proposed to implement sampling for \gls{ids}:  \cite{estan_new_2003} propose to single out large flows for \gls{ids} and sample fewer packets for increased computational efficiency. \cite{bakhoum_intrusion_2011} use an approach based on Markov chains to sample packets for \gls{ids}.
\cite{murali_kodialam_detecting_2003} lay out a game-theoretic model for determining which network paths are more vulnerable and require more packet sampling. 
\cite{ha_suspicious_2016} develop a system for Software Defined Networks, which aims to sample more packets from more vulnerable parts of the network.

Regarding \gls{rnn} for \gls{ids}, \cite{hartl_explainability_2019} build an LSTM-based recurrent classifier for Network Intrusion Detection and develop methods for explainability and to assess the security of recurrent classifiers.

Some works investigated ways to let \glspl{rnn} skip parts of sequences. The focus of these works was mostly Natural Language Processing and we know of no work that focuses on network traffic. The following papers provide an overview of this research domain:

\cite{yu_learning_2017} propose using \gls{rl} to let an LSTM learn to skim text. Their technique does not aim to maximize sparsity but only lets the \gls{rl} optimize classification performance for the sequence as a whole. Thus, if it learns to skip elements it is only because it is better for achieving good classification performance, not because it wants to increase sparsity.

\cite{campos_skip_2018} develop a technique that includes a \textit{skip gate} into LSTM and other recurrent cells. It works without \gls{rl} but on the downside, their method can only be trained once and not be adapted afterwards and also it explicitly depends on the the implementation of the underlying classifier, while we want a solution in which the sampling procedure and the classifier are independent.

\cite{seo_neural_2018} use an LSTM with one full state vector and one reduced one. At each step they decide whether to use the reduced or the full state vector and can reduce computational cost by using the reduced one. For network traffic we think that it is often not even necessary to have a reduced state vector since for some attack flows, already the first couple of packets might contain enough information and then the remaining packets of the flow do not have to be considered anymore at all.

\cite{gui_long_2018} use \gls{rl} but their intention is not to skip samples but instead to make the \gls{rl} learn to choose a previous state that can aid at the current step. Thus it tries to help the \gls{rnn} with memorizing information from the past but does not actually take less data, which is the goal we pursue in this work.

These proposed techniques focus their evaluation on text and hence do not fulfill all criteria we set out in \autoref{sec:introduction}. Furthermore, all techniques that use \gls{rl} use discrete actions, meaning that there is a hyperparameter $k$, which influences the maximum jump that is possible. However, for a network flow, already after the second packet it might be obvious that a flow is an attack and thus the remaining packets can be completely ignored. It is thus beneficial to not have a fixed maximum step size but a continuous one. This allows skipping arbitrarily many packets.

%\section{Setup}
%%ci

\section{\ours{}}

%Commonly, network packet sampling is performed on backbone infrastructures where, due to large bandwidths, inspecting every packet in detail is expensive. Sampling is thus required to save computational power and expenses or even because current \gls{ml} and specifically \gls{dl} methods might even require more computational power than can be delivered by present-day hardware. Assuming that the behavior of the network follows a predefined underlying process, sampling is therefore not detrimental on average to statistical analysis or measurement. Although the dataset in this work does not come from a backbone network, sampling reduces the high computational effort due to additional processing required by the complex \gls{ml}-based \glspl{ids}.
To fulfill the goals outlined in \autoref{sec:introduction}, in this section we propose the architecture of our proposed sampling framework \ours.

\subsection{Classifier}
\label{subsec:classifier}
We implemented a three-layer LSTM-based classifier with 128 neurons at each layer. We choose 128 neurons as we don't notice a significant degradation of performance compared to the results of \cite{hartl_explainability_2019} (check Table II, flow accuracy) and as fewer neurons allowed for more agility during the training phase and allowed us to run more experiments. As the input features we use source port, destination port, protocol identifier, packet length, \gls{iat} to the previous packet in the flow, packet direction (i.e. forward or reverse path) and all TCP flags (0 if the flow is not TCP).
We omitted \gls{ttl} values, as they are likely to lead to unwanted prediction behaviour \cite{bachl_walling_2019}.  Among the used features, source port, destination port and protocol identifier are constant over the whole flow while the others vary.
Additionally, we add the number of skipped packets since the last packet as a feature, which can help the classifier because otherwise it might be confused because of missing packets.
We used the usual 5-tuple flow key, which distinguishes flows based on the protocol they use and their source and destination port and IP address.

\subsection{Reinforcement Learning-based Sampling}
\label{subsec:rlForSampling}

Given enough training instances, \gls{rl} algorithms have been shown to perform very well in complex decision-making situations such as playing video games \cite{mnih_playing_2013} as well as networking problems such as congestion control \cite{bachl_rax_2019}. We thus consider an \gls{rl}-based architecture to be suitable for taking packet sampling decisions. The system should look at the current packet as well as the history of the conversation flow and output how many packets should be skipped until the next packet is considered in the current flow. The reward that the \gls{rl} tries to maximize should be a combination of classification accuracy and the number of packets that could be skipped (sparsity). An operator should be able to specify how much accuracy is allowed to be traded off for achieving higher sparsity.% and the objective of the \gls{rl} at a timestep t of a flow is defined as:

%\begin{align}
%\begin{split}
%U_t = & \text{future average classification performance} \\
%+ & \alpha\times\text{future average sparsity}
%\end{split}
%\end{align}

For a flow of length $N$ at each packet $n$ we have two reward metrics which we compute in a fashion that is known from \textit{R-learning} \cite{schwartz_reinforcement_1993}:

%\begin{figure}
%\centering
%  \includegraphics[width=\columnwidth]{img/rewards.pdf}
%  \caption{The computation of the rewards for a sample flow: Gray packets are skipped while white ones are chosen. At the top row the resulting rewards for classification for each chosen packet are depicted and in the bottom row the ones for sparsity. Numbers are rounded to two digits after the decimal point.}
%  \label{fig:rewards}
%\end{figure}

\begin{figure*}
\centering
  \includegraphics[width=0.8\textwidth]{img/rewards_calc_example.pdf}
  \caption{The computation of the rewards for a sample malicious flow: Gray packets (0) are skipped while black ones (1) are chosen. ``\#Skipped\//\#Total'' (number of packets skipped with respect to the total number of packets both starting from the current position) and ``Confidence'' are the underlying parameters for rewards calculations (see \autoref{eq:R_classification} and \autoref{eq:R_sparsity}). Numbers are rounded to two digits after the decimal point.}
  \label{fig:rewards}
\end{figure*}

The classification reward metric (\autoref{eq:R_classification}) captures how correct the classifier is on average for all future packets of a flow. The label of each packet in a flow has a value of 0 for non-attack traffic and 1 for attack traffic with one flow being only attack or non-attack but not mixed. The confidence is the sigmoided $\left(\frac{e^x}{1+e^x}\right)$ output of the classifier, yielding a value between 0 and 1, where 0 is absolute confidence for a flow being benign and 1 is absolute confidence for a flow being an attack.
\begin{align}
\begin{split}
R_{\text{classification},n} = \frac{1}{N-n} \sum_{i=n+1}^{N} |1 - \left(\text{label}_i - \text{confidence}_i\right)|
\end{split}
\label{eq:R_classification}
\end{align}

Intuitively, the classification reward makes sure, that packets are chosen that contain the maximum accuracy and thus result in the highest classification performance.

The sparsity reward metric (\autoref{eq:R_sparsity}) captures how much sparsity is achieved on average for all future packets of a flow. E.g. for a flow of length 10 if after the 5th packet, 3 packets are chosen and 2 are skipped from what is left, the sparsity reward metric at packet 5 is $\frac{2}{10-(4+1)}=\frac{2}{5}$.
%\note{blue}{Fares: This equation can be shortened by diving by the total count of packets left.\\
%Fares: If the length of the flow is not known, how do we estimate T}.
%\note{red}{Max: It's only relevant during training when you know the lengths of the flows. In deployment, you don't need the rewards anymore...}
\begin{align}
\begin{split}
R_{\text{sparsity},n} = \frac{\sum_{i=n+1}^{N} 1_{\text{skipped},i}}{N-(n+1)}
\end{split}
\label{eq:R_sparsity}
\end{align}

\autoref{fig:rewards} shows how classification and sparsity rewards are computed for a toy example.

We opt for an \emph{\gls{ac}} approach (as shown in \autoref{fig:sampling}) modeled after \cite{mnih_asynchronous_2016} because we consider it being more interpretable than a Deep \gls{rl} approach based on a \textit{Deep Q Network} like \cite{mnih_playing_2013}, as it outputs the reward that was expected and the reward that was actually achieved, which we explain in more detail in the next paragraphs.

Our proposed framework (see \autoref{fig:neuralNetworkArchitecture}) consists of three independent neural networks: the classifier, the critic and the actor.

The classifier's goal is to guess correctly if a flow is an attack or not. For this, it outputs its confidence for non-attack/attack after reading a packet. As stated in \autoref{subsec:classifier}, we choose an LSTM-based supervised classifier for this work, but our framework can work together with any classifier as long as it outputs a confidence for non-attack/attack at each packet and can handle skipped packets. For instance it would also be possible to replace the LSTM with an unsupervised autoencoder akin to \cite{mirsky_kitsune_2018}.

The critic aims to estimate the future expected average classification performance and the future expected average sparsity that can be achieved at a packet $n$. For this, given the current input vector and the state of the LSTM cells $s_n$ and the neural network weights of the critic $\theta_v$, it outputs the value functions $v_{\text{classification},n}$ and $v_{\text{sparsity},n}$ at each packet $n$. It aims to predict the classification performance and sparsity correctly for each packet by minimizing the following loss function:
\begin{align}
\begin{split}
l_{\text{v},n} =& \left(R_{\text{classification},n} - v_{\text{classification},n}\left(s_n ; \theta_\text{c} \right)\right)^2 \\
+ & \left(R_{\text{sparsity},n} - v_{\text{sparsity},n}\left(s_n ; \theta_\text{c} \right)\right)^2
\end{split}
\end{align}

Now that the reward metrics and the critic's outputs are defined, we can define the overall utility $U$ that we want to maximize at each packet $n$ as follows:
\begin{align}
\begin{split}
U_n =& \left(R_{\text{classification},n} + \alpha \cdot R_{\text{sparsity},n}\right) \\
- & \left(v_{\text{classification},n} + \alpha \cdot v_{\text{sparsity},n}\right)
\end{split}
\end{align}

With $\alpha$ being the sparsity-accuracy tradeoff parameter. This means that we want the reward to be higher than the critic's expectation. Note that it would be also possible to unify both reward metrics in one and have the critic have only one output, but we consider it to be more interpretable to have two separate outputs, which can be inspected by an operator.

The actor outputs a probability distribution at packet $n$, which is sampled to determine the number of packets that should be skipped next. It aims to change the distribution so that actions which result in a higher reward are chosen more frequently, maximizing the utility function. However, this strategy could lead to the actor getting stuck with what it considers the best decision and never try alternatives even though they might result in a higher reward. This would be equivalent to the actor being stuck in a local maximum. Thus, the actor not only tries to change the distribution so that the actions become more optimal (higher reward) but at the same time also aims to maximize the entropy of the distribution so that alternative choices are still explored reasonably often.

Specifically, at packet $n$ of a network flow, given the current state $s_n$ (which consists of the current input as well as the LSTM state) and the neural network weights of the actor $\theta_a$, the actor outputs a probability distribution $\pi \left( a_n \given s_n ; \theta_\text{a} \right)$ from which the action $a_n \ge 1$ is sampled. $a_n=1$ means that the next packet is inspected next, while, for example, $a_n=2$ means that one packet is skipped until the next that is chosen, $a_n = 3$ means that two are skipped and so forth. Besides wanting to make actions which result in a high reward more likely, the actor also seeks to optimize the entropy $H(\cdot)$ of the probability distribution so that it keeps exploring and doesn't get stuck with a suboptimal policy.

Combining these objectives, the actor network aims to minimize the loss function
\begin{align}
\begin{split}
l_{\text{a},n} =& \underbrace{ -\log \left( \pi \left( a_n \given s_n ; \theta_\text{a} \right) \right) U_n}_{\text{Policy Loss}} \\
&- \beta \underbrace{H\left( \pi\left( s_n; \theta_a \right)\right) }_{\text{Entropy}}
\end{split}\label{eq:actor}
\end{align}

%The Critic network evaluates the behavior of the Actor network and computes the classification performance and sparsity that it thinks is going to be achieved.% using the computed value function and gives it feedback based on the performance obtained by the LSTM classifier.
%The number of packets to skip is taken from the Actor's output distribution and fed into the packet parser, which is responsible for counting the packets and selecting the appropriate one.

In related work regarding \gls{rl}-based sampling for sequences, only discrete actions were evaluated so far: The actor would output a categorical probability distribution with a fixed maximum number of bins $k$. A drawback of this approach is that only a predefined maximum number of packets can be skipped at once (the maximum of the actions space) and that for a very large number of $k$, it can take a long time to converge to an optimum policy as there are $k$ different options to explore at each packet $n$. Besides discrete actions, we thus also experiment with continuous actions, which allow to skip arbitrarily many packets with long-tailed distributions.

\begin{figure}[h]
\centering
  \includegraphics[width=\columnwidth]{neural_network.pdf}
  \caption{An overview of the complete neural network being used in our experiments. Ocher stands for the input, blue for linear layers, red for LSTM layers and purple for the outputs of the neural network. The numbers in parentheses next to each label stand for the width of that layer (i.e.~the number of neurons (except for the inputs and outputs)). Parts of the neural network are shared between $v_\text{classification}$ and $v_\text{sparsity}$ as well as $\mu$ and $\sigma$ as they output quantities of similar magnitude and thus the gradients that are computed when updating the neural network are of the same order of magnitude as well. Our implementation also supports sharing the LSTM between all three neural network. However, as the five outputs can have different orders of magnitude, larger gradients would ``overwrite'' the gradients of the smaller ones.}
  \label{fig:neuralNetworkArchitecture}
\end{figure}

%\note{blue}{Fares: can be moved one section above\\}
%An overview of the neural networks of the classifier, actor and critic that we use in this work is given in \autoref{fig:neuralNetworkArchitecture}.

\begin{figure}[h]
\centering
  \includegraphics[width=0.8\columnwidth]{img/rnn-sampling.pdf}
  \caption{SparseIDS framework. Three deep networks are shown: the LSTM classifier, the actor network and the critic network. The actor takes actions based on the feedback from the the critic network and the classifier. The critic network tracks the actor's actions based on the classifier confidence and finally, the packets selector parses packets based on the decision of the actor network.}
  \label{fig:sampling}
\end{figure}

%\note{blue}{Fares: can be removed, it was already mentioned.\\}
%A schematic overview over the \gls{rl} for sampling is given in \autoref{fig:sampling}.

\subsection{Continuous Actions}

For performing continious actions we need a probability distribution which can be parametrized to have the mean between 0 and positive infinity. Furthermore it must have at least two parameters so that the standard deviation, which is required for the entropy (\autoref{eq:actor}), can be changed independently of the mean. For example, the Exponential distribution can be parametrized to have its mean anywhere between 0 and positive infinity, however, it has only one parameter and thus the mean and the standard deviation cannot be independently changed. Conversely, a simple distribution which fits our requirements is the log-normal distribution, which is defined as a normal distribution whose output is exponentiated:

%Let $Z$ be a random variable from a standard normal distribution $\mathcal{N}(0, 1)$ and let $\mu_{\text{normal}}$ and $\sigma_{\text{normal}} > 0$ be two real numbers. Then
%\begin{align}
%X = e^{\mu_{\text{normal}}+\sigma_{\text{normal}} Z}
%\end{align}
%is a log-normally distributed variable. Note that the log-normal distribution's mean $\mu$ and its standard deviation $\sigma$ are different from $\mu_\text{normal}$ and $\sigma_\text{normal}$, which are used to parametrize the distribution.

The architecture in \autoref{fig:neuralNetworkArchitecture} shows that the actor network has two outputs: $\mu$ and $\sigma$. These encode the mean and standard deviation of a log-normal distribution and can never be smaller or equal to zero and hence we always apply a rectifier function at the output. We choose the softplus, which is a smooth rectifier function:
\begin{align}
\text{softplus}(x) = \log\left(1 + e^x \right)
\end{align}

When experimenting with continuous actions, we encountered one practical problem: At the beginning of training, the classifier would still perform badly and would not give the actor a useful reward. As a result, the actor would only focus on sparsity and aim to maximize it, as the classification reward would always be equally bad no matter how many packets are chosen. Thus, the actor would learn to let $\mu$ go to infinity. This would increase the reward at the beginning when the interaction between the three components of our \gls{rl} system is still fragile. As a solution, we add a penalty to the reward calculation when the actor jumps far behind the last flow packet. The modified reward applies only to the last chosen packet and penalizes the actor jumping more than one packet behind the last packet. The sparsity reward formula (\autoref{eq:R_sparsity}) of the last packet is modified as follows, where \textit{last} refers to the index of the last chosen packet in the flow ($n=\textit{last}$):

\begin{align}
\begin{split}
R_{\text{sparsity},\textit{last}} = \frac{\left(\sum_{i=\textit{last}+1}^{N} 1_{\text{skipped},i}\right)}{N- (\textit{last}+1+\left(\textit{last}+a_{\textit{last}} - N - 1\right))}
\end{split}
\label{eq:R_sparsity_last}
\end{align}

Specifically, it adds the term $\left(\textit{last}+a_{\textit{last}} - N - 1\right)$ to the denominator which penalizes the actor jumping more than one packet behind the flow. Jumping just right after the flow does not get penalized because if the actor considers the rest of the flow as irrelevant, ignoring the rest of the flow and jumping to the first packet behind the flow is a useful action.

\section{Other sampling techniques}
\label{sec:other_sampling_techniques}
In order to evaluate the performance of \ours, we compare it with a few straightforward sampling techniques.
We opt for three sampling families that, unlike \ours{} cannot adapt their strategy based on packet contents. % (i.e. sampling rates are constant and not correlated with subsequent analysis): %The underlying principle of each family is summarized as follows:
% Max: Don't understand

\subsection{Random Sampling}

Given the fraction $p$ of packets to be selected in a sequence of $N$ packets, this technique will sample each individual packet with probability of $p$. Theoretically, if the sequence is large enough (infinitely large), $m$ packets will be randomly sampled from a uniform distribution whereby $m \approx \lfloor N \times p \rfloor$. Nevertheless, in the case of short flows, $m$ is not constant and varies to some degree.  In practice, these technique works well if several packets have similar properties (redundancy) and therefore the probability of skipping important information is minimal.

\subsection{Relative First \emph{m} Packets}
As its name suggests, this technique only samples the first packets from each flow and ignores the rest of the flow. For a sampling rate $p$ it will take the first $m$ packets from each flow so that $m = \lfloor N \times p \rfloor$. For example, for a flow with length 10 and $p=0.2$, it will take the first 2 packets, for another flow of length 40 and for the same sampling rate ($p=0.2$), it will sample the first 8 packets. For this procedure the length of a flow has to be known beforehand and thus it is not applicable in a streaming scenario. This technique is successful under the assumption that most information for determining maliciousness is present in the first few packets of a flow, i.e. a large data transfer in which the initial packets represent a hand-shake after which the majority of packets contain the raw data and do not reveal significant information to non-\gls{dpi} systems.

\subsection{First \emph{m} Packets}
This technique is similar to the one above, with the difference that the length of a flow doesn't have to be known beforehand. If flows have an average length of 10 packets, and the sampling rate $p=0.2$ it would take up to two packets from each flow regardless of its length.

\subsection{Every \emph{i}th Sampling}
Finally we use a periodic sampling technique that for a sampling rate $p$, it takes a sample each $i= \lfloor \frac{1}{p} \rfloor$ packet(s). Similar to the last technique, the length of the flow is not needed in this case. In practice, this technique is best suited to keep the distribution of the sampled set similar to the original distribution, and it is thus suited for scenarios where information is heterogeneously distributed between packets over an entire network flow.

In each of the above techniques, the first packet of each flow is always taken because otherwise one could get undefined behavior: If a flow only consists of 1 packet and this packet is not chosen, it is not even defined if the prediction of the classifier was correct for this flow or not. Thus, in our implementations of the above sampling techniques, we make sure that even when the first packet is always chosen, the overall fraction of chosen packets corresponds to the sampling probability $p$.

After applying the sampling techniques, the sequence of selected packets is then fed to the classifier discussed in \autoref{subsec:classifier}, where features are extracted and a decision is made after each packet. Also, the classifier gets information on how many packets were skipped prior to each packet.

% Max: If we write things like this they might think we were too lazy to implement them :D
%There exist many more rule-based sampling techniques e.g., temporal, adaptive etc. However, studying their behaviors is out of the scope of this paper.

\section{Dataset}

%\note{blue}{Fares: if we submit to CNS (page limit is 9), we can remove this sections as it is not relevant and statistics can be found online}
% Max: Should be fine :)

\begin{table}[h]
\caption{Occurrence frequency of the attack flows of CIC-IDS-2017.}
\label{tab:occurrence}
\centering
\begin{tabular}{l r}
\toprule
Attack type & \hspace*{-4mm}Proportion \\ \midrule
DoS Hulk	&	10.10\%	\\
PortScan, Firewall	&	6.90\%	\\
DDoS LOIT	&	4.08\%	\\
Infiltration	&	3.30\%	\\
DoS GoldenEye	&	0.32\%	\\
DoS SlowHTTPTest	&	0.18\%	\\
DoS Slowloris	&	0.17\%	\\
Brute-force SSH	&	0.11\%	\\
Botnet ARES	&	0.03\%	\\
XSS attack	&	0.03\%	\\
PortScan, no Fw.	&	0.02\%	\\
Brute-force FTP	&	0.01\%	\\
SQL injection	&	$<$0.01\%	\\
Heartbleed	&	$<$0.01\%	\\
\bottomrule
\end{tabular}
\end{table}

%\begin{table}[h]
%\caption{Occurrence frequency of the attack flows.}
%\label{tab:occurrence}
%\centering
%\subfloat[CIC-IDS-2017\label{fig:cicids17proportions}
%]{
%\begin{tabular}{l r}
%\toprule
%Attack type & \hspace*{-4mm}Proportion \\ \midrule
%DoS Hulk	&	10.10\%	\\
%PortScan, Firewall	&	6.90\%	\\
%DDoS LOIT	&	4.08\%	\\
%Infiltration	&	3.30\%	\\
%DoS GoldenEye	&	0.32\%	\\
%DoS SlowHTTPTest	&	0.18\%	\\
%DoS Slowloris	&	0.17\%	\\
%Brute-force SSH	&	0.11\%	\\
%Botnet ARES	&	0.03\%	\\
%XSS attack	&	0.03\%	\\
%PortScan, no Fw.	&	0.02\%	\\
%Brute-force FTP	&	0.01\%	\\
%SQL injection	&	$<$0.01\%	\\
%Heartbleed	&	$<$0.01\%	\\
%\bottomrule
%\end{tabular}
%}{}
%\subfloat[UNSW-NB15\label{fig:unswnb15proportions}
%]{
%\begin{tabular}{l r}
%\toprule
%Attack type & \hspace*{-4mm}Proportion \\ \midrule
%Exploits	&	1.42\%	\\
%Fuzzers	&	1.01\%	\\
%Reconnaissance	&	0.58\%	\\
%Generic	&	0.21\%	\\
%DoS	&	0.19\%	\\
%Shellcode	&	0.08\%	\\
%Analysis	&	0.03\%	\\
%Backdoors	&	0.02\%	\\
%Worms	&	0.01\%	\\
%\bottomrule
%\end{tabular}
%}{}
%\end{table}

For our experiments, we use the \textit{CIC-IDS-2017} \cite{sharafaldin_toward_2018} dataset, which includes more than 2 million flows of network data, containing both benign traffic and a large number of different attacks. The distribution of different attacks in the datasets are shown in \autoref{tab:occurrence}.

We use Z-score normalization and a train/test split of 2:1. To speed up experimentation, we cut flows at a maximum length of 20. As Figure \ref{fig:inspectingAll} shows, there are only few flows that exceed that length.

\section{Results}

\begin{figure}[h]
\centering
  \includegraphics[width=\columnwidth]{{"img/events.out.tfevents.1579352256.gpu"}.pdf}
  \caption{Training with a tradeoff of 0.1 with continuous actions. Training accuracy increases sharply in the beginning while becoming slower later on. Sparsity increases sharply in the beginning and then more slowly. As the classifier gets better, the \gls{rl} starts skipping more packets.}
  \label{fig:training}
\end{figure}

%The results of the sampling part are discussed in this section.
We study the effect of the sparsity-accuracy tradeoff first, then we evaluate both discrete and continuous actions and, finally, we compare our results with other sampling techniques. Overall, sampling based on \gls{rl} reveals that a high number of packets can be skipped while maintaining good accuracy. Then, we inspect how the \gls{rl} chooses packets and if some patterns can be spotted.

For all our experiments we train our networks for at least 8 epochs ($>10$ million flows for both datasets) with the Adam optimizer with a learning rate of $0.001$. An important implementation detail is that we always start with the first packet of a flow, meaning that it cannot be skipped. Also, we set the entropy parameter to $0.01$. 
\autoref{fig:training} shows how, during training, accuracy increases quickly at the beginning and then gradually afterwards. At the same time, sparsity gets higher as the \gls{rl} learns which packets can be skipped since they do not influence classification performance. %We then evaluate the resulting neural networks on the test dataset.

Opposed to the behavior during training, we do not randomly sample from the probability distribution of the actor during deployment. We instead take the mean of the distribution for the continuous case and the mode for the discrete one. This is reasonable and best practice \cite{mnih_asynchronous_2016} since during deployment we do not aim to explore the possible choices anymore as they were already learned during training: Only the optimal choice, which was learned, should be applied and no exploration is necessary since the actor does not train anymore. Furthermore, the critic is only necessary during training for the reward of the actor. During deployment, it can be disabled, saving computational power.

%\begin{table}
%\caption{Comparing classification metrics per flow with different tradeoffs for both CIC-IDS-2017 and UNSW-NB15.} \label{tab:results_tradeoff}
%\centering
%\begin{tabular*}{\columnwidth}{>{\rowmac}l @{\extracolsep{\fill}} >{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c<{\clearrow}} \toprule
%& \multicolumn{3}{c}{CIC-IDS-2017} & \multicolumn{3}{c}{UNSW-NB15} \\
%Tradeoff & 0.0 & 0.1 & 0.5 & 0.0 & 0.1 & 0.5 \\	\midrule
%\setrow{\bfseries}Sparsity & 0\% & 76.3\% & 76.5\% & 0\% & 83.5\% & 84.5\% \\ \midrule
%\setrow{\bfseries}Accuracy & 99.5\% & 99.4\% & 99.3\% & 98.9\% & 98.4\% & 98.5\% \\
%Precision & 99.4\% & 98.5\% & 98.7\% & 81.1\% & 69.5\% & 70.8\% \\
%Recall & 98.5\% & 99.0\% & 98.6\% & 84.0\% & 89.9\% & 89.6\% \\
%F1 & 99.0\% & 98.7\% & 98.6\% & 82.5\% & 78.4\% & 79.1\% \\
%Youden & 98.4\% & 98.5\% & 98.2\% & 83.3\% & 88.6\% & 88.4\% \\
%\bottomrule
%\end{tabular*}
%\end{table}

\begin{table}
\caption{Comparing classification metrics per flow with different tradeoffs.} \label{tab:results_tradeoff}
\centering
\begin{tabular}{>{\rowmac}l >{\rowmac}r>{\rowmac}r>{\rowmac}r<{\clearrow}} \toprule
Tradeoff & 0.0 & 0.1 & 0.5 \\	\midrule
\setrow{\bfseries}Sparsity & 0\% & 76.3\% & 76.5\% \\ \midrule
\setrow{\bfseries}Accuracy & 99.5\% & 99.4\% & 99.3\% \\
Precision & 99.4\% & 98.5\% & 98.7\% \\
Recall & 98.5\% & 99.0\% & 98.6\% \\
F1 & 99.0\% & 98.7\% & 98.6\% \\
Youden & 98.4\% & 98.5\% & 98.2\% \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Influence of the tradeoff}

\autoref{tab:results_tradeoff} shows the influence of the tradeoff parameter $\alpha$ for both datasets. For columns with a tradeoff of 0.0, we train the classifier regularly with all packets without using \gls{rl}. For tradeoffs 0.1 and 0.5 we train \ours{} using \gls{rl} with continuous actions and keep the tradeoff constant for the whole training. Results show that the higher the tradeoff, the less packets are sampled and the lower the accuracy.

\subsection{Discrete vs.~Continuous actions}

%\begin{table}
%\caption{Comparing classification metrics per flow with either discrete (20 different actions) or continuous actions for CIC-IDS-2017 and UNSW-NB15. The tradeoff is always $0.1$.} \label{tab:results_disc_cont}
%\centering
%\begin{tabular*}{\columnwidth}{>{\rowmac}l @{\extracolsep{\fill}} >{\rowmac}c>{\rowmac}c>{\rowmac}c>{\rowmac}c<{\clearrow}} \toprule
%& \multicolumn{2}{c}{CIC-IDS-2017} & \multicolumn{2}{c}{UNSW-NB15} \\
%Actions & cont. & disc. & cont. & disc. \\	\midrule
%\setrow{\bfseries}Sparsity & 76.3\% & 73.6\% & 83.5\% & 86.6\% \\ \midrule
%\setrow{\bfseries}Accuracy & 99.4\% & 99.2\% & 98.4\% & 98.4\% \\
%Precision & 98.5\% & 98.7\% & 69.5\% & 68.0\% \\
%Recall & 99.0\% & 98.3\% & 89.9\% & 94.6\% \\
%F1 & 99.7\% & 98.5\% & 78.4\% & 79.2\% \\
%Youden & 98.5\% & 97.8\% & 88.6\% & 93.2\% \\
%\bottomrule
%\end{tabular*}
%\end{table}

\begin{table}
\caption{Comparing classification metrics per flow with either discrete (20 different actions) or continuous actions. The tradeoff is always $0.1$.} \label{tab:results_disc_cont}
\centering
\begin{tabular}{>{\rowmac}l >{\rowmac}r>{\rowmac}r<{\clearrow}} \toprule
Actions & continuous & discrete \\	\midrule
\setrow{\bfseries}Sparsity & 76.3\% & 73.6\% \\ \midrule
\setrow{\bfseries}Accuracy & 99.4\% & 99.2\% \\
Precision & 98.5\% & 98.7\% \\
Recall & 99.0\% & 98.3\% \\
F1 & 99.7\% & 98.5\% \\
Youden & 98.5\% & 97.8\% \\
\bottomrule
\end{tabular}
\end{table}

\autoref{tab:results_disc_cont} compares continuous and discrete (20 different) actions and shows that continuous actions result in a higher accuracy with fewer samples being selected. This leads to the conclusion, that at least for our use case and for the dataset we consider, there is no consistent difference in performance for continuous and discrete actions. Generally, the advantage of discrete actions is that they are theoretically more powerful as they can compare each of the possible actions with each other, while the only possibility for continuous actions is to increase or decrease the mean of the log-normal distribution, which can lead to worse performance. On the other hand, with discrete actions, the maximum number of samples to skip is fixed, which we consider to be a major drawback. Furthermore, training time is potentially higher since, for example, for a discrete action space of 20 actions, each option is treated as being completely different and, while actually there is not a large difference between skipping 18 or 19 packets, there is a big difference between skipping 0 packets compared to skipping 19 packets. This means that for discrete actions, all actions are perceived as being completely independent, which is not true as the reward function considers sparsity, which increases gradually as more packets are skipped and thus similar actions result in similar rewards.

\subsection{Comparing with other sampling methods}

\begin{table}
\caption{Comparing classification metrics per flow between \gls{rl}-based sampling with continuous actions (tradeoff $0.1$) and alternative strategies.} \label{tab:results_others}
\centering
\begin{tabular}{>{\rowmac}l >{\rowmac}r>{\rowmac}r>{\rowmac}r>{\rowmac}r>{\rowmac}r<{\clearrow}} \toprule
& \gls{rl} & random & every $i$th & first $n$ & first $n$ equal \\	\midrule
\setrow{\bfseries}Sampled & 76.3\% & 76.3\% & 76.3\% & 76.3\% & 76.3\% \\ \midrule
\setrow{\bfseries}Accuracy & 99.4\% & 96.9\% & 97.8\% & 97.3\% & 98.3\% \\
Precision & 98.5\% & 92.4\% & 95.8\% & 93.8\% & 95.6\% \\
Recall & 99.0\% & 95.6\% & 95.4\% & 95.5\% & 97.9\% \\
F1 & 99.7\% & 94.0\% & 95.6\% & 94.7\% & 96.7\% \\
Youden & 98.5\% & 93.0\% & 93.9\% & 93.4\% & 96.4\% \\
\bottomrule
\end{tabular}
\end{table}

Next we compared the \gls{rl} technique used by \ours{} with several intuitive sampling strategies suggested in \autoref{sec:other_sampling_techniques}.
\autoref{tab:results_others} reveals that \ours{} achieves the highest accuracy with the same sparsity. Not surprisingly, random sampling performs the worst, as it is likely that a flow is ``unfortunate'' and doesn't get enough packets sampled while another one gets more packets or the packets that carry more useful information. The best strategy besides the \gls{rl} approach seems to be ``first $n$ equal'' sampling, which matches the conclusion from \cite{hartl_explainability_2019} that the first couple of packets are the most important ones for determining maliciousness of a flow.
``first $n$ equal'' performs better than ``first $n$'' because for ``first $n$'', longer flows will have more packets sampled from them. Conversely, very short flows will get few packets. For example, with $p=0.2$, a flow of length 20 will get 4 packets while one with 4 packets will get only 1. However, with `first $n$ equal'', the short and the long flow will get the same number of packets. Apparently it is better if all flows get the same number of packets than if long flows get more packets.

\subsection{Inspecting the \gls{rl}'s behavior}
\label{subsec:inspecting}

\begin{figure*}[h]
\centering
\subfloat[Benign Traffic\label{fig:inspectingNormal}
]{
\includegraphics[width=0.98\columnwidth]{{"../plots/plot_rl/flows_Jan18_14-06-48_gpu_lstm_module_8.pth_prediction_outcomes_0_3.pickle_10_Normal"}.pdf}
}{}
\subfloat[SSH Patator\label{fig:inspectingPatator}
]{
\includegraphics[width=0.98\columnwidth]{{"../plots/plot_rl/flows_Jan18_14-06-48_gpu_lstm_module_8.pth_prediction_outcomes_0_3.pickle_2_Brute Force-SSH-Patator"}.pdf}
}{}
\subfloat[DDoS Slowloris\label{fig:inspectingSlowloris}
]{
\includegraphics[width=0.98\columnwidth]{{"../plots/plot_rl/flows_Jan18_14-06-48_gpu_lstm_module_8.pth_prediction_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
}{}
\subfloat[All samples\label{fig:inspectingAll}
]{
\includegraphics[width=0.98\columnwidth]{{"../plots/plot_rl/flows_Jan18_14-06-48_gpu_lstm_module_8.pth_prediction_outcomes_0_3.pickle_15_All samples"}.pdf}
}{}
\caption{Inspecting the sampling strategy of \ours{} for various attack types with continuous actions and a tradeoff of $0.1$. The blue line shows the confidence. The gray bars show how many samples had this number of packets. For example, the gray bar in Figure \ref{fig:inspectingNormal} at position 19 with height 10000 signifies that 10000 flows of benign traffic had a length of at least 20 packets (since we say that the first packet is at position 0). The red bars represent the count of packets sampled at this position for all flows. The first packet is always chosen for each flow by design.}
\label{fig:inspecting}
\end{figure*}

\autoref{fig:inspecting} shows the sampling strategy for different attack types. Figure \ref{fig:inspectingNormal} shows that for benign traffic, the sampling strategy is quite heterogeneous: Not specific packets get always chosen, but different packets get chosen, depending on the flow. This contrasts with SSH Patator (Figure \ref{fig:inspectingPatator}), where the 0th, the 3rd, the 13th and the 15th packet are always chosen, since they appear to be especially indicative of an attack. Also, the flows seem to contain no useful information between the 3rd and the 13th packet, so that this part is always skipped altogether. Slowloris (Figure \ref{fig:inspectingSlowloris}) is more similar to what we saw for benign traffic: The sampling strategy varies a lot for different flows and not always the same pattern is used. When looking at all flows (Figure \ref{fig:inspectingAll}) it becomes apparent that the second packet (index 1) is almost never chosen, and the third is also chosen quite rarely. We attribute this to the fact that most communication flows are TCP and use handshakes, meaning that the second packet is simply a reply a part of the handshake, revealing no important information. In general, we see that the sampling strategy varies significantly between different attack types and for some attack types even for different flows within the same attack type, showing that \ours{} learns to choose the packets effectively, which reveal the malicious nature of a flow.

One concern could be the overly deterministic nature of the decisions of \ours{} for some attack types: An adversary could see that only certain packets of a flow are looked at by the \gls{rl} system and thus only manipulate the few packets that the classifier looks at to generate adversarial samples with methods as studied by \cite{hartl_explainability_2019}. One potential remedy for this is to also randomly sample from the distribution of the actor during deployment like it is done during training. This could result in slightly worse accuracy but lead to behavior that is harder to predict for an adversary. Currently we take the mean of the probability distribution for continuous actions and the mode for the discrete one, resulting in deterministic behavior.

\section{Steering}

So far, we have always trained \ours{} with a fixed tradeoff parameter $\alpha$. Nevertheless, there is also an important use case in which an operator wants a certain minimum sparsity, for example 50\%, to limit the necessary computation power and thus reduce costs.
We thus propose a method that enables steering the tradeoff to achieve a certain desired sparsity. The steering algorithm works as follows:
\begin{enumerate}
\item A model is trained for which the tradeoff is chosen randomly for each flow during training from a uniform distribution yielding values between 0 and 1. We add a new feature to each packet that indicates the tradeoff chosen for the respective flow so that it can adjust its sampling strategy in anticipation of the tradeoff that will determine the reward.
\item During deployment, the tradeoff is set to the maximum that was used during training (1 in our case), which results in maximum sparsity. The tradeoff is then continuously decremented if the current sparsity is still above the specified minimum. It is important that the tradeoff can never fall below 0 because this is a case that has not been encountered during training and can lead to abnormal behavior.
\end{enumerate}

\begin{figure}[h]
\centering
  \includegraphics[width=\columnwidth]{{"img/lstm_module_8.pth_global_tradeoff_1.0_steering_step_size_0.05_steering_target_sparsity_0.5_batches_to_consider_for_steering_500_sampling_rl"}.pdf}
  \caption{Steering the tradeoff with an allowed minimum sparsity of 50\%. The tradeoff is initialized as 1 and continuously lowered as long as the sparsity is higher than 50\%. The procedure stops if the minimum sparsity is reached or the tradeoff reached its minimum of 0, which is the case here. Sparsity falls from around 74.5\% to 73.5\%.}
  \label{fig:steering}
\end{figure}

%\autoref{fig:steering} shows the steering in action. It stops when the tradeoff reaches 0 and the specified minimum sparsity of 50\% has not been reached yet. It might seem surprising that even with a parameter of 0, which means that only classification matters and sparsity not at all, still a sparsity of 73.5\% is achieved. We attribute this to the fact that \ours{} learns that some flows contain may contain packets that do not contain any useful information about determining maliciousness. Thus they are even skipped when there is no reward for skipping them because they do not increase the classification score at all.

\autoref{fig:steering} shows the steering in action. We start with a maximum sparsity of ~74.4\% and try to reach 50\%. To do so, the steering system starts decreasing the tradeoff parameter until it reaches 0. It might seem surprising that even with a tradeoff of 0, which means that only classification matters and not sparsity not at all, a sparsity of 73.5\% is still achieved. We attribute this to the fact that \ours{} learns that some flows may contain packets that do not contribute in determining maliciousness and are skipped since they do not affect the accuracy. %\item the actor still in any case attempts to maximize the entropy of its distribution which means that when no packets are skipped, entropy is very low. \end{enumerate*}%Nothworthy is the fact that the goal sparsity (50\%) is not reached due to the previous assumptions, however, a decreasing slope is clearly shown in figure 4

\section{Conclusion}
\label{sec:conclusion}
In this paper we aimed to develop smart packet sampling methods for the use case of \glspl{ids} to save computational power. To this end, we proposed \ours{} which is a framework for packet sampling based on \gls{rl}.

% Max: 1/4 is already pronounced one fourth
We show that the \gls{rl}-based approach with a tradeoff between classification accuracy and sparsity indeed results in higher sparsity the larger the tradeoff parameter, while potentially classification performance is sacrificed. Surprisingly, we saw that \ours{} on average chooses very few packets per flow: Typically $\approx \frac{1}{4}$. Furthermore, our results showed that \gls{rl}-based sampling can intelligently choose relevant packets for intrusion detection and outperform non-smart sampling techniques.

Besides discrete actions for our \gls{rl} system we also implemented continuous ones. We showed that a naive implementation would not work as the training is too fragile in the beginning and the \gls{rl} system learned to maximize only one of the rewards (sparsity) in the beginning. This was solved by adding a penalty for jumping behind the end of the sequence which proved effective. Comparing with discrete actions, continuous actions seem to perform on par. This is still an advantage, as for continuous actions the neural network only needs to have two outputs (mean and standard deviation) while for discrete ones it needs as many outputs as there are actions. The continuous action framework is thus more lightweight and also can learn to skip arbitrarily many packets.

Inspecting the decisions of our \gls{rl} framework showed that the sampling strategy is clearly different for different attack strategies and also within some attack families, indicating that \ours{} indeed adopts different sampling strategies for different conversation flows.

Eventually, we implement a steering method to help achieving a certain sparsity threshold. The steering mechanism can be used to limit the computational expenses of an operator.

In general, we think that \gls{rl}-based packet sampling is not only applicable to \gls{ids} but we are confident that other domains of networking can benefit from similar solutions to save computational expenses.

\section*{Acknowledgements}
The Titan Xp used for this research was donated by the NVIDIA Corporation.

\renewcommand*{\bibfont}{\small}
\bibliographystyle{ieeetr}
\bibliography{bibliography}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.